---
title: "Andropogon_virginicus - Global Model Outputs"
author: "Kelsey Brock"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---
```{r}
starttime <- Sys.time()
```

# Name of Species being modelled
```{r}
spname <- "Andropogon_virginicus"
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,error=TRUE,message=FALSE)
```
```{r}
#get the needed packages
if(!require("pacman")){
	install.packages("pacman")
	library(pacman)}
p_load("dplyr", "PresenceAbsence", "DAAG", "ggplot2", 'tidyr', "terra", "sf", "knitr", "dismo", "gam", "randomForest", "gbm", "biomod2", "CoordinateCleaner")
```
```{r}
# setting paths
  path.root <- "~/codingwork/fire"  # typical class root dir

  path.in.general <- paste(path.root, "/input", sep = "")
  
  path.in.specific <- paste(path.root, "/input/", spname, sep = "")
  path.out.specific <- paste(path.root, "/output/", spname, sep = "") 
  path.out.specific.ensemble.global <- paste(path.root, "/output/", spname, "/", spname, "_ensemble-projection_global", sep = "")
  path.out.specific.ensemble.global.current.HIdata <- paste(path.root, "/output/", spname, "/", spname, "_ensemble-projection_GLOBAL_current", sep = "")
  path.out.specific.ensemble.global.NCAR8.5 <- paste(path.root, "/output/", spname, "/", spname, "_ensemble-projection_GLOBAL_NCAR8.5", sep = "")
  path.out.specific.ensemble.global.IPRC8.5 <- paste(path.root, "/output/", spname, "/", spname, "_ensemble-projection_GLOBAL_IPRC8.5", sep = "")
  
  path.out.specific.ensemble.local <- paste(path.root, "/output/", spname, "/", spname, "_ensemble-projection_local", sep = "")
```

```{r}
# handy projection string
prj.wgs84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"  # epsg:4326
```

####  Get the Occurrence Data

```{r}
setwd(path.in.specific)
global.df <- read.csv(paste(spname, "_tr_global_allvars.csv", sep = ""), header=T, sep=',', stringsAsFactors=F)
head(global.df) # look at data - should include all variables
```

```{r}
table(global.df$scientificName) # examine  frequencies
```

### Get selected variables during data exploration phase
```{r}
setwd(path.in.specific)
keep <- get(load(file = paste0( spname, "_list_picked_vars_global.RData")))
keep
```

### Import raster of current and future climate scenarios

#### Current


```{r}
setwd(path.in.specific)
#rasterOptions(maxmemory=1e+07)
#stack <- get(load("bioclimstack.RData"))
stack <- terra::rast( paste0(spname, "_bioclimstack.tif"))
stack
```


```{r}
stack <- terra::subset(stack, keep)
#stack3 <- stack(stack)
stack

#writeRaster(stack3,"tempstack.img", format = "HFA", overwrite = TRUE)
```

# BIOMOD2

```{r}
global.df
```

```{r}
responsevar <- c("FNETID", "scientificName", "tr.Lon", "tr.Lat" )
responsevar
keep
```

```{r}
global.df <- global.df %>% dplyr::select(responsevar, keep) 
global.df
```


Getting a weird error!?!?  this is probably the chunk that's the problem.  Double check indexing
```{r}
# Clean data
myrespvar <- as.numeric(global.df$scientificName)
myRespName <- spname
myexplvar <- global.df[, 5:length(global.df)]
myresp.xy <- global.df[, 3:4]
clean_data <- BIOMOD_FormatingData(resp.var=myrespvar, expl.var=myexplvar, resp.xy=myresp.xy, resp.name=myRespName)
```

```{r}
clean_data
```

```{r}
my_models <- c("MAXENT","GLM","GAM","RF","GBM", "MARS", "FDA") # Algorithms to fit the models
#my_models <- c("GLM","GAM","MARS", "FDA") # Algorithms to fit the models
```

```{r}
# Settings for the model algorithms
model_opt <- BIOMOD_ModelingOptions(GLM=list(type='quadratic', interaction.level=0),GBM=list(n.trees=1000), MAXENT = list(path_to_maxent.jar = "C:/RFolder/maxent.jar",product=FALSE),GAM=list(k=3))

```

MAKE SURE YOU CHANGE # of reps to 10!!
```{r}
all_models <- BIOMOD_Modeling(bm.format = clean_data,
                                    modeling.id = 'global',
                                    models = my_models,
                                    bm.options = model_opt,
                                    CV.strategy = 'random',
                                    CV.nb.rep = 10,
                                    CV.perc = 0.80,
                                    metric.eval = c('TSS','ROC', 'KAPPA'),
                                    seed.val = 4)
```

```{r}
all_models
```

```{r}
# get all models evaluation
myBiomodModelEval <- get_evaluations(all_models)
myBiomodModelEval
```

```{r}
Evals <- bm_PlotEvalMean(all_models, group.by = "algo",
                    metric.eval = c("ROC","TSS"))
Evals$plot
```
```{r}
eval_table <- Evals$tab[,1:4] %>% dplyr::rename("ROCmean" = "mean1", "TSSmean" = "mean2")
eval_table
setwd(path.out.specific)
write.csv(eval_table, file = paste0( spname, "eval_table_allalgos_global.csv"))
```


```{r}

mods <- get_built_models(all_models, algo = "GLM" )
bm_PlotResponseCurves(
      all_models,
      models.chosen = mods,
      new.env = get_formal_data(all_models, "expl.var"),
      show.variables = get_formal_data(all_models, "expl.var.names"),
      fixed.var = "mean",
      do.bivariate = FALSE,
      do.plot = TRUE,
      do.progress = TRUE)

```


### Project to HI


```{r}
xmin = -160.5613
xmax =  -154.7061
ymin = 17.9578 
ymax =  22.3496

projPoly <- terra::as.polygons(terra::ext(xmin, xmax, ymin, ymax))

projPoly
```


```{r}
setwd(path.in.general)
baseline <- list.files(path = path.in.general, pattern = "wc2.1_30s", full.names = T) # list of .img files; $ strips extra
combined <- terra::rast(baseline)# examine


names(combined) <- c("bio1",  "bio12", "bio15")

combined <- terra::subset(combined, keep)
#stack3 <- stack(stack)
combined 
```

```{r}
library(raster)
predsProj <- terra::crop(combined, projPoly )
#predsProj <- raster::mask(predsProj, projPoly)
#predsProj <- raster::stack(predsProj)
predsProj
```
```{r}
library(raster)
predsProj <- raster::stack(predsProj)
  predsProj
```
```{r}
clean_data
```


```{r}
# projection over Hawaii under current conditions
myBiomodProj <- BIOMOD_Projection(
bm.mod = all_models,
new.env = predsProj,
proj.name = 'current',
selected.models = 'all',
binary.meth = 'TSS',
compress = 'xz',
clamping.mask = F,
output.format = '.img')
myBiomodProj
```

```{r}
# make some plots sub-selected by str.grep argument
#plot(myBiomodProj, str.grep = 'GLM')
```


```{r}
myBiomodEM <- BIOMOD_EnsembleModeling(
bm.mod = all_models,
models.chosen = 'all',
em.by='all',
metric.select = c('ROC'),
metric.select.thresh = c(0.65),
metric.eval = c('ROC', 'TSS', "KAPPA"),
em.algo = c('EMmean', 'EMcv', 'EMci', 'EMmedian', 'EMca', 'EMwmean'),
committee.averaging = T,
prob.mean.weight = T,
EMwmean = 'proportional' )
```
```{r}
Evals <- bm_PlotEvalMean(myBiomodEM, group.by = "algo",
                    metric.eval = c("ROC","TSS"))
Evals$plot
```
```{r}
eval_table <- Evals$tab[,1:4] %>% dplyr::rename("ROCmean" = "mean1", "TSSmean" = "mean2")
eval_table
setwd(path.out.specific)
write.csv(eval_table, file = paste0( spname, "eval_table_ensemble_global.csv"))
```




```{r}
myBiomodEF_current <- BIOMOD_EnsembleForecasting(
bm.em = myBiomodEM,
bm.proj = myBiomodProj)
myBiomodEF_current
```

```{r}
#plot(myBiomodEF_current, str.grep = 'EMcaByROC')
```
```{r}
get_predictions(myBiomodEF_current)
```


```{r}
## assign ensemble forecasting object with a new name for notation
## convenience
setwd(path.out.specific)
ensproj <- myBiomodEF_current

## get the rasterstack of ensemble projections
ensproj.stack <- get_predictions(ensproj)

## define the output directory
outdir.ensproj <- paste0(spname, '_ensemble-projection_global')
dir.create(outdir.ensproj, showWarnings = FALSE, recursive = TRUE)

## save individual layers on the hard drive
purrr:::map(
names(ensproj.stack),
~ writeRaster(
ensproj.stack[[.x]],
overwrite = TRUE,
filename = file.path(outdir.ensproj, paste0(.x, '.tiff'))
)
)

list.files('outdir.ensproj')
```

```{r}
setwd(path.out.specific.ensemble.global)
final_raster_global <- raster("Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo.tiff")

final_raster_global

plot(final_raster_global)
```





## Projecting to Hawaii using the HI-specific climate data (Only needed for comparison phase)

### Import raster of current and future climate scenarios

#### Current

```{r}
setwd(path.in.general)
stack <- terra::rast( "HI_bioclimstack.tif")

xmin = -160.3722
xmax =  -154.5714
ymin =  18.7084
ymax =  22.4615

local.bufbbox <- extent(xmin, xmax, ymin, ymax)


stack <- crop(stack, local.bufbbox )
stack
```

```{r}
# projection over the globe under current conditions
myBiomodProj_currentHIdata <- BIOMOD_Projection(
bm.mod = all_models,
new.env = stack,
proj.name = 'current-HIdata',
selected.models = 'all',
binary.meth = 'TSS',
compress = 'xz',
clamping.mask = F,
output.format = '.img')
myBiomodProj_currentHIdata
```

```{r}
myBiomodEF_currentHIdata <- BIOMOD_EnsembleForecasting(
bm.em = myBiomodEM,
bm.proj = myBiomodProj_currentHIdata)
myBiomodEF_currentHIdata
```

```{r}
## assign ensemble forecasting object with a new name for notation
## convenience
setwd(path.out.specific)
ensproj <- myBiomodEF_currentHIdata

## get the rasterstack of ensemble projections
ensproj.stack <- get_predictions(ensproj)

## define the output directory
outdir.ensproj <- paste0(spname, '_ensemble-projection_GLOBAL_current')
dir.create(outdir.ensproj, showWarnings = FALSE, recursive = TRUE)

## save individual layers on the hard drive
purrr:::map(
names(ensproj.stack),
~ writeRaster(
ensproj.stack[[.x]],
overwrite = TRUE,
filename = file.path(outdir.ensproj, paste0(.x, '.tiff'))
)
)

list.files('outdir.ensproj')
```

```{r}
setwd(path.out.specific.ensemble.global.current.HIdata)
final_raster_global <- raster("Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo.tiff")

final_raster_global

plot(final_raster_global)
```

#### FUTURE

#### Future NCAR RCP 8.5

```{r}
setwd(path.in.general)
NCAR8.5 <- terra::rast("new_NCAR_future_projection_rcp85_bioclims.tif")
NCAR8.5
```
```{r}
seasonal <- list.files(path = path.in.general, pattern = "rcp85_NCAR", full.names = T)
seasonal

seasonal_stack <- terra::rast(seasonal)# examine
seasonal_stack
```

```{r}
NCAR8.5_stack <- c(NCAR8.5, seasonal_stack)
NCAR8.5_stack
```
    
```{r}
names(NCAR8.5_stack) <- c("bio1",  "bio2",  "bio3",  "bio4",  "bio5",  "bio6",  "bio7",  "bio8",  "bio9", "bio10", "bio11", "bio12", "bio13", "bio14", "bio15", "bio16", "bio17", "bio18", "bio19", "bio20", "bio21", "bio22", "bio23", "bio24", "bio25", "bio26", "bio27"  )
names(NCAR8.5_stack)
NCAR8.5_stack 
```
### Get selected variables during data exploration phase
```{r}
setwd(path.in.specific)
keep <- get(load(file = paste0( spname, "_list_picked_vars_global.RData")))
keep
```
```{r}
NCAR8.5_stack <- terra::subset(NCAR8.5_stack, keep)
#stack3 <- stack(stack)
NCAR8.5_stack 
```

```{r}
# projection over the globe under current conditions
myBiomodProj_NCAR8.5HIdata <- BIOMOD_Projection(
bm.mod = all_models,
new.env = NCAR8.5_stack,
proj.name = 'future-NCAR8.5',
selected.models = 'all',
binary.meth = 'TSS',
compress = 'xz',
clamping.mask = F,
output.format = '.img')
myBiomodProj_NCAR8.5HIdata
```

```{r}
myBiomodEF_futureNCAR8.5HIdata <- BIOMOD_EnsembleForecasting(
bm.em = myBiomodEM,
bm.proj = myBiomodProj_NCAR8.5HIdata)
myBiomodEF_futureNCAR8.5HIdata
```

```{r}
## assign ensemble forecasting object with a new name for notation
## convenience
setwd(path.out.specific)
ensproj <- myBiomodEF_futureNCAR8.5HIdata

## get the rasterstack of ensemble projections
ensproj.stack <- get_predictions(ensproj)

## define the output directory
outdir.ensproj <- paste0(spname, '_ensemble-projection_GLOBAL_NCAR8.5')
dir.create(outdir.ensproj, showWarnings = FALSE, recursive = TRUE)

## save individual layers on the hard drive
purrr:::map(
names(ensproj.stack),
~ writeRaster(
ensproj.stack[[.x]],
overwrite = TRUE,
filename = file.path(outdir.ensproj, paste0(.x, '.tiff'))
)
)

list.files('outdir.ensproj')
```

```{r}
setwd(path.out.specific.ensemble.global.NCAR8.5)
final_raster_global <- raster("Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo.tiff")

final_raster_global

plot(final_raster_global)
```



#### Future IPRC RCP 8.5

```{r}
setwd(path.in.general)
IPRC8.5 <- terra::rast("new_IPRC_future_projection_rcp85_bioclims.tif")
IPRC8.5
```

```{r}
seasonal <- list.files(path = path.in.general, pattern = "rcp85_IPRC", full.names = T)
seasonal

seasonal_stack <- terra::rast(seasonal)# examine
seasonal_stack
```

```{r}
IPRC8.5_stack <- c(IPRC8.5, seasonal_stack)
IPRC8.5_stack
```
    
```{r}
names(IPRC8.5_stack) <- c("bio1",  "bio2",  "bio3",  "bio4",  "bio5",  "bio6",  "bio7",  "bio8",  "bio9", "bio10", "bio11", "bio12", "bio13", "bio14", "bio15", "bio16", "bio17", "bio18", "bio19", "bio20", "bio21", "bio22", "bio23", "bio24", "bio25", "bio26", "bio27" )
names(IPRC8.5_stack)

IPRC8.5_stack 
```
### Get selected variables during data exploration phase
```{r}
setwd(path.in.specific)
keep <- get(load(file = paste0( spname, "_list_picked_vars_global.RData")))
keep
```
```{r}
IPRC8.5_stack <- terra::subset(IPRC8.5_stack, keep)
#stack3 <- stack(stack)
IPRC8.5_stack 
```

```{r}
# projection over the globe under current conditions
myBiomodProj_IPRC8.5HIdata <- BIOMOD_Projection(
bm.mod = all_models,
new.env = IPRC8.5_stack,
proj.name = 'future-IPRC8.5',
selected.models = 'all',
binary.meth = 'TSS',
compress = 'xz',
clamping.mask = F,
output.format = '.img')
myBiomodProj_IPRC8.5HIdata
```

```{r}
myBiomodEF_IPRC8.5HIdata <- BIOMOD_EnsembleForecasting(
bm.em = myBiomodEM,
bm.proj = myBiomodProj_IPRC8.5HIdata)
myBiomodEF_IPRC8.5HIdata
```

```{r}
## assign ensemble forecasting object with a new name for notation
## convenience
setwd(path.out.specific)
ensproj <- myBiomodEF_IPRC8.5HIdata

## get the rasterstack of ensemble projections
ensproj.stack <- get_predictions(ensproj)

## define the output directory
outdir.ensproj <- paste0(spname, '_ensemble-projection_GLOBAL_IPRC8.5')
dir.create(outdir.ensproj, showWarnings = FALSE, recursive = TRUE)

## save individual layers on the hard drive
purrr:::map(
names(ensproj.stack),
~ writeRaster(
ensproj.stack[[.x]],
overwrite = TRUE,
filename = file.path(outdir.ensproj, paste0(.x, '.tiff'))
)
)

list.files('outdir.ensproj')
```

```{r}
setwd(path.out.specific.ensemble.global.IPRC8.5)
final_raster_global <- raster("Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo.tiff")

final_raster_global

plot(final_raster_global)
```



```{r}
nowtime <- Sys.time()
elapsed = nowtime - starttime
print(elapsed)
```


<!-- ```{r} -->
<!-- newdir <- paste(path.out.specific, "/", outdir.ensproj,  sep = "") -->
<!-- setwd(newdir) -->
<!-- final_raster <- raster("Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo.tiff") -->

<!-- final_raster -->

<!-- plot(final_raster) -->
<!-- ``` -->



<!-- Need to standardize the probabilities (from 0to 1) -->

<!-- ```{r} -->
<!-- maxValue(final_raster) -->
<!-- minValue(final_raster) -->
<!-- ``` -->

<!-- ```{r} -->

<!--   layers <- {} # initialize (empty) list of raster layers -->
<!--   for (i in 1:length(names(final_raster))) { -->
<!--     m1 <- final_raster[[i]] # get a prob map -->
<!--     m2 <- 1/maxValue(final_raster[[i]]) * final_raster[[i]] # standardize all probs to max=1 -->
<!--     m3 <- unlist(strsplit(names(final_raster[[i]]), "[.]")) # split prob layer name apart -->
<!--     names(m2) <- paste(m3[1], ".", m3[2], sep = "")  # assign name to raster value -->
<!--     assign(paste(m3[1], ".", m3[2], sep = ""), m2) # assign new name to standardized layer -->
<!--     layers <- c(layers, get(paste(m3[1], ".", m3[2], sep = ""))) -->
<!--   } -->
<!--   final_raster <- stack(layers) -->
<!--   final_raster -->
<!-- ``` -->
<!-- ```{r} -->
<!-- plot(final_raster) -->
<!-- maxValue(final_raster) -->
<!-- minValue(final_raster) -->
<!-- ``` -->

<!-- uugh. can't get this to work for some reason -->
<!-- ```{r} -->
<!-- terra::rast(final_raster) -->
<!-- setwd(path.out.specific) -->
<!-- raster::writeRaster(final_raster, filename = paste0( spname,"_final_global_ensemble_raster.img"), overwrite = T) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- setwd(path.out.specific) -->
<!-- # Save an object to a file -->
<!-- saveRDS(final_raster, file = paste0( spname,"_final_global_ens_rast.RDS")) -->
<!-- ``` -->
