---
title: "Holcus lanatus - local Model Outputs"
author: "Kelsey Brock"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r}
starttime <- Sys.time()
```
# Name of Species being modelled
```{r}
spname <- "Andropogon_virginicus"
```
```{r}
futureclim <- "RCP8.5"
fclim <- 85
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,error=TRUE,message=FALSE)
```
```{r}
#get the needed packages
if(!require("pacman")){
	install.packages("pacman")
	library(pacman)}
p_load("dplyr", "PresenceAbsence", "DAAG", "ggplot2", 'tidyr', "raster", "sf", "knitr", "dismo", "gam", "randomForest", "gbm", "biomod2")
```
```{r}
# setting paths
  path.root <- "~/codingwork/fire"  # typical class root dir
  path.in.general <- paste(path.root, "/input", sep = "")
  path.in.specific <- paste(path.root, "/input/", spname, sep = "")
  path.out.specific <- paste(path.root, "/output/", spname, sep = "")
  path.out.specific.ensemble.global <- paste(path.root, "/output/", spname, "/", spname, "_ensemble-projection_global", sep = "")
  path.out.specific.ensemble.local <- paste(path.root, "/output/", spname, "/", spname, "_ensemble-projection_local", sep = "")
```
```{r}
# handy projection string
prj.wgs84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"  # epsg:4326
```

####  Get the Occurrence Data

```{r}
setwd(path.in.specific)
local.df <- read.csv(paste(spname, "_tr_local_allvars.csv", sep = ""), header=T, sep=',', stringsAsFactors=F)
local.df # look at data - should include all variables
```
```{r}
table(local.df$scientificName) # examine  frequencies
```
### Get Global Projection for Pseudoabsence down-weighting


```{r}
setwd(path.out.specific.ensemble.global)
GlobalProj <- raster("Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo.tiff")

GlobalProj

plot(GlobalProj)
```


Need to standardize the probabilities (from 0to 1)

```{r}
maxValue(GlobalProj)
minValue(GlobalProj)
```

```{r}

  layers <- {} # initialize (empty) list of raster layers
  for (i in 1:length(names(GlobalProj))) {
    m1 <- GlobalProj[[i]] # get a prob map
    m2 <- 1/maxValue(GlobalProj[[i]]) * GlobalProj[[i]] # standardize all probs to max=1
    m3 <- unlist(strsplit(names(GlobalProj[[i]]), "[.]")) # split prob layer name apart
    names(m2) <- paste(m3[1], ".", m3[2], sep = "")  # assign name to raster value
    assign(paste(m3[1], ".", m3[2], sep = ""), m2) # assign new name to standardized layer
    layers <- c(layers, get(paste(m3[1], ".", m3[2], sep = "")))
  }
  GlobalProj <- stack(layers)
  GlobalProj
```
```{r}
maxValue(GlobalProj)
minValue(GlobalProj)
```

If you get "too large for int" warning, I think it can be ignored
Straight out of the publiction ---
```{r}
weights_gl <- 1/(1+((GlobalProj/(GlobalProj - 1))^2)) # inverse logistic transformation

weights_gl
```


```{r}
xmin = -160.3722
xmax =  -154.5714
ymin =  18.7084
ymax =  22.4615

local.bufbbox <- extent(xmin, xmax, ymin, ymax)

weights_gl <- raster::crop(weights_gl, local.bufbbox )

weights_gl
```

```{r}
pres <- subset(local.df, scientificName == 1)
pres$yweights <- 1

abs <- subset(local.df, scientificName == 0)
```
```{r}
weights_gl
```




```{r}
weight_values_ab <- raster::extract(weights_gl, cbind(abs$decimalLongitude, abs$decimalLatitude))

abs$yweights <- weight_values_ab
```
```{r}
local.df <- rbind(pres, abs)
local.df
```
```{r}
max(local.df$yweights)
min(local.df$yweights)
mean(local.df$yweights)
```
```{r}
absences <- subset(local.df, local.df$scientificName == 0)


hist(absences$yweights)
```




### Get selected variables during data exploration phase
```{r}
setwd(path.in.specific)
keep <- get(load(file = paste0( spname, "_list_picked_vars_local.RData")))
keep
```

```{r}
header <- c("FNETID", "scientificName", "decimalLongitude", "decimalLatitude")
oo <- c(header, keep)

picked.df <- local.df %>% dplyr::select(oo)
picked.df
```

```{r}
pres <- subset(picked.df, scientificName == 1)
pres$yweights <- 1

abs <- subset(picked.df, scientificName == 0)
```

```{r}
weight_values_ab <- raster::extract(weights_gl, cbind(abs$decimalLongitude, abs$decimalLatitude))

abs$yweights <- weight_values_ab
```
```{r}
picked.df <- rbind(pres, abs)
picked.df
```

### Import raster of current and future climate scenarios

#### Current

```{r}
setwd(path.in.general)
stack <- terra::rast( "HI_bioclimstack.tif")

xmin = -160.3722
xmax =  -154.5714
ymin =  18.7084
ymax =  22.4615

local.bufbbox <- extent(xmin, xmax, ymin, ymax)


stack <- crop(stack, local.bufbbox )
stack

```

#### FUTURE


<!-- ```{r} -->
<!-- setwd(path.in.general) -->
<!-- #baseline <- list.files(pattern = "new_baseline_bioclims.tif") # list of .img files; $ strips extra -->
<!-- #baseline_stack <- stack(baseline)# examine -->
<!-- elevation <- terra::rast("elevhi.tif") -->
<!-- elevation <- terra::project(elevation, stack) -->
<!-- elevation -->
<!-- ``` -->




#### Future NCAR RCP 8.5

```{r}
setwd(path.in.general)
NCAR8.5 <- terra::rast("new_NCAR_future_projection_rcp85_bioclims.tif")
NCAR8.5
```

```{r}
seasonal <- list.files(path = path.in.general, pattern = "rcp85_NCAR", full.names = T)
seasonal

seasonal_stack <- terra::rast(seasonal)# examine
seasonal_stack
```

```{r}
NCAR8.5_stack <- c(NCAR8.5, seasonal_stack)
NCAR8.5_stack
```
    
```{r}
names(NCAR8.5_stack) <- c("bio1",  "bio2",  "bio3",  "bio4",  "bio5",  "bio6",  "bio7",  "bio8",  "bio9", "bio10", "bio11", "bio12", "bio13", "bio14", "bio15", "bio16", "bio17", "bio18", "bio19", "bio20", "bio21", "bio22", "bio23", "bio24", "bio25", "bio26", "bio27"  )
names(NCAR8.5_stack)

NCAR8.5_stack 
```
#### Future IPRC RCP 8.5

```{r}
setwd(path.in.general)
IPRC8.5 <- terra::rast("new_IPRC_future_projection_rcp85_bioclims.tif")
IPRC8.5
```

```{r}
seasonal <- list.files(path = path.in.general, pattern = "rcp85_IPRC", full.names = T)
seasonal

seasonal_stack <- terra::rast(seasonal)# examine
seasonal_stack
```

```{r}
IPRC8.5_stack <- c(IPRC8.5, seasonal_stack)
IPRC8.5_stack
```
    
```{r}
names(IPRC8.5_stack) <- c("bio1",  "bio2",  "bio3",  "bio4",  "bio5",  "bio6",  "bio7",  "bio8",  "bio9", "bio10", "bio11", "bio12", "bio13", "bio14", "bio15", "bio16", "bio17", "bio18", "bio19", "bio20", "bio21", "bio22", "bio23", "bio24", "bio25", "bio26", "bio27" )
names(IPRC8.5_stack)

IPRC8.5_stack 
```


```{r}
#cmip_eightfive <- raster::getData(name = 'CMIP5', var = 'bio', res = 2.5,
#        rcp = fclim, model = 'AC', year = 70,
#        path = paste0(path.in.general, "/bioclims_CMIP5_ssp585"))
```

```{r}
# get WorldClim bioclimatic variable rasters
#elev <- raster::getData(name = "worldclim", var = "alt", res = 2.5, lat = , lon = )
#elev
```

```{r}
#path.preds <- paste(path.root, "/input/yearly_and_seasonal_projections_HI",  sep = "")
##setwd(path.preds)
#future <- list.files(pattern = "rcp85.tif") # list of .img files; $ strips extra
#cmip_eightfive <- stack(future)# examine
#cmip_eightfive
```

```{r}
#cmip_eightfive <- raster::stack(cmip_eightfive, elev)
#names(cmip_eightfive)
```

<!-- ```{r} -->
<!-- names(cmip_eightfive) <- names(stack) -->

<!-- xmin = -160.3722 -->
<!-- xmax =  -154.5714 -->
<!-- ymin =  18.7084 -->
<!-- ymax =  22.4615 -->

<!-- local.bufbbox <- extent(xmin, xmax, ymin, ymax) -->


<!-- cmip_eightfive <- crop(cmip_eightfive, local.bufbbox ) -->

<!-- cmip_eightfive -->
<!-- ``` -->

# BIOMOD2 

```{r}
#my_models <- c("MAXENT","GLM","GAM","RF","GBM", "MARS", "FDA") # Algorithms to fit the models
my_models <- c("MAXENT","GLM","GAM","MARS", "FDA") # Algorithms to fit the models
```

```{r}
# Settings for the model algorithms
model_opt <- BIOMOD_ModelingOptions(GLM=list(type='quadratic', interaction.level=0),GBM=list(n.trees=1000), MAXENT = list(path_to_maxent.jar = "C:/RFolder/maxent.jar",product=FALSE),GAM=list(k=3))

```


```{r}
responsevar <- c("FNETID", "scientificName", "decimalLongitude", "decimalLatitude" )
responsevar
keep
yweights <- c("yweights")
yweights
```

```{r}
local.df <- local.df %>% dplyr::select(responsevar, keep, yweights) 
local.df
```

```{r}
# Clean data
myrespvar <- as.numeric(local.df$scientificName)
myRespName <- spname
myexplvar <- local.df[, 5:(length(local.df)-1)]
myresp.xy <- local.df[, 3:4]
clean_data <- BIOMOD_FormatingData(resp.var=myrespvar, expl.var=myexplvar, resp.xy=myresp.xy, resp.name=myRespName)
```
```{r}
clean_data
```

```{r}
all_models <- BIOMOD_Modeling(bm.format = clean_data,
                                    modeling.id = 'local',
                                    models = my_models,
                                    bm.options = model_opt,
                                    weights=local.df$yweights,
                                    CV.strategy = 'random',
                                    CV.nb.rep = 10,
                                    CV.perc = 0.8,
                                    metric.eval = c('TSS','ROC', 'KAPPA'),
                                    seed.val = 42)

```
```{r}
all_models
```
```{r}
# get all models evaluation
myBiomodModelEval <- get_evaluations(all_models)
# print the dimnames of this object
dimnames(myBiomodModelEval)
myBiomodModelEval
```


```{r}
Evals <- bm_PlotEvalMean(all_models, group.by = "algo",
                    metric.eval = c("ROC","TSS"))
Evals$plot
```
```{r}
eval_table <- Evals$tab[,1:4] %>% dplyr::rename("ROCmean" = "mean1", "TSSmean" = "mean2")
eval_table
setwd(path.out.specific)
write.csv(eval_table, file = paste0( spname, "eval_table_allalgos_local.csv"))
```



```{r}

mods <- get_built_models(all_models, algo = "GAM" )
bm_PlotResponseCurves(
      all_models,
      models.chosen = mods,
      new.env = get_formal_data(all_models, "expl.var"),
      show.variables = get_formal_data(all_models, "expl.var.names"),
      fixed.var = "mean",
      do.bivariate = FALSE,
      do.plot = TRUE,
      do.progress = TRUE)

```


### Project to HI


```{r}
xmin = -160.5613
xmax =  -154.7061
ymin = 17.9578 
ymax =  22.3496

projPoly <- terra::as.polygons(terra::ext(xmin, xmax, ymin, ymax))

projPoly
```
```{r}
stack <- terra::subset(stack, keep)
#stack3 <- stack(stack)
stack 
```

```{r}
predsProj <- terra::crop(stack, projPoly )
#predsProj <- raster::mask(predsProj, projPoly)
#predsProj <- raster::stack(predsProj)
predsProj
```
# ```{r}
# library(raster)
# predsProj <- raster::stack(predsProj)
#   predsProj
# ```


<!-- ```{r} -->
<!-- xmin = -160.5613 -->
<!-- xmax =  -154.7061 -->
<!-- ymin = 17.9578  -->
<!-- ymax =  22.3496 -->

<!-- projPoly <- terra::as.polygons(terra::ext(xmin, xmax, ymin, ymax)) -->

<!-- projPoly -->
<!-- ``` -->


```{r}
stack3 <- stack(stack)
stack3
```

```{r}
# projection over the globe under current conditions
myBiomodProj <- BIOMOD_Projection(
bm.mod = all_models,
new.env = stack3,
proj.name = 'current',
selected.models = 'all',
binary.meth = 'TSS',
compress = 'xz',
clamping.mask = F,
output.format = '.img')
myBiomodProj

```

```{r}
# make some plots sub-selected by str.grep argument
#plot(myBiomodProj, str.grep = 'GLM')
```


```{r}
myBiomodEM <- BIOMOD_EnsembleModeling(
bm.mod = all_models,
models.chosen = 'all',
em.by='all',
metric.select = c('ROC'),
metric.select.thresh = c(0.40),
metric.eval = c('ROC', 'TSS', "KAPPA"),
em.algo = c('EMmean', 'EMcv', 'EMci', 'EMmedian', 'EMca', 'EMwmean'),
committee.averaging = T,
prob.mean.weight = T,
EMwmean = 'proportional' )

```
```{r}
# get all models evaluation
myBiomodModelEval <- get_evaluations(myBiomodEM)
myBiomodModelEval
```
```{r}
Evals <- bm_PlotEvalMean(myBiomodEM, group.by = "algo",
                    metric.eval = c("ROC","TSS"))
Evals$plot
```
```{r}
eval_table <- Evals$tab[,1:4] %>% dplyr::rename("ROCmean" = "mean1", "TSSmean" = "mean2")
eval_table
setwd(path.out.specific)
write.csv(eval_table, file = paste0( spname, "eval_table_ensemble_local.csv"))
```


```{r}
myBiomodEF_current <- BIOMOD_EnsembleForecasting(
bm.em = myBiomodEM,
bm.proj = myBiomodProj)
myBiomodEF_current

```
```{r}
get_predictions(myBiomodEF_current)
```


```{r}
#plot(myBiomodEF_current, str.grep = 'Cenchrus.setaceus_EMwmeanByROC_mergedAlgo_mergedRun_mergedData')
```


```{r}
## assign ensemble forecasting object with a new name for notation
## convenience
setwd(path.out.specific)
ensproj <- myBiomodEF_current

## get the rasterstack of ensemble projections
ensproj.stack <- get_predictions(ensproj)

## define the output directory
outdir.ensproj <- paste0(spname, '_ensemble-projection_local')
dir.create(outdir.ensproj, showWarnings = FALSE, recursive = TRUE)

## save individual layers on the hard drive
purrr:::map(
names(ensproj.stack),
~ writeRaster(
ensproj.stack[[.x]],
overwrite = TRUE,
filename = file.path(outdir.ensproj, paste0(.x, '.tiff'))
)
)

list.files('outdir.ensproj')
```


```{r}
setwd(path.out.specific.ensemble.local)
final_raster <- raster("Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo.tiff")

final_raster

plot(final_raster)
```


###save
```{r}
setwd(path.out.specific)
terra::writeRaster(final_raster, paste0(spname,"_current", "_local", "_final_ensemble_raster_local.img"),  overwrite = TRUE)
```

## FUTURE CLIMATE NCAR8.5

```{r}
NCAR8.5_stack
```
```{r}

NCAR8.5_stack <- terra::subset(NCAR8.5_stack, keep)
#stack3 <- stack(stack)
NCAR8.5_stack 
```
```{r}
predsProjNCAR <- terra::crop(NCAR8.5_stack, projPoly )
#predsProj <- raster::mask(predsProj, projPoly)
predsProjNCAR <- raster::stack(predsProjNCAR)
predsProjNCAR
```

```{r}
# projection over the globe under future conditions
myBiomodProj_NCAR <- BIOMOD_Projection(
bm.mod = all_models,
new.env = predsProjNCAR,
proj.name = 'future_NCAR',
selected.models = 'all',
binary.meth = 'TSS',
compress = 'xz',
clamping.mask = F,
output.format = '.img')
myBiomodProj_NCAR

```


```{r}
# make some plots sub-selected by str.grep argument
#plot(myBiomodProj_NCAR, str.grep = 'GLM')
```


```{r}
myBiomodEF_futureNCAR <- BIOMOD_EnsembleForecasting(
bm.em = myBiomodEM,
bm.proj = myBiomodProj_NCAR)
myBiomodEF_futureNCAR
```


```{r}
## assign ensemble forecasting object with a new name for notation
## convenience
setwd(path.out.specific)
ensproj <- myBiomodEF_futureNCAR

## get the rasterstack of ensemble projections
ensproj.stack <- get_predictions(ensproj)

## define the output directory
outdir.ensproj <- paste0(spname, '_ensemble-projection_local_NCAR8.5')
dir.create(outdir.ensproj, showWarnings = FALSE, recursive = TRUE)

## save individual layers on the hard drive
purrr:::map(
names(ensproj.stack),
~ writeRaster(
ensproj.stack[[.x]],
overwrite = TRUE,
filename = file.path(outdir.ensproj, paste0(.x, '.tiff'))
)
)

list.files('outdir.ensproj')
```


```{r}
newdir <- paste(path.out.specific, "/", outdir.ensproj,  sep = "")
setwd(newdir)
final_raster_NCAR8.5 <- raster("Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo.tiff")

final_raster_NCAR8.5

plot(final_raster_NCAR8.5)
```
###save
```{r}
setwd(path.out.specific)
terra::writeRaster(final_raster_NCAR8.5, paste0(spname,"_NCAR8.5", "_local", "_final_ensemble_raster.img"),  overwrite = TRUE)
```


## FUTURE CLIMATE IPRC8.5

```{r}
IPRC8.5_stack
```
```{r}

IPRC8.5_stack <- terra::subset(IPRC8.5_stack, keep)
#stack3 <- stack(stack)
IPRC8.5_stack 
```
```{r}
predsProjIPRC <- terra::crop(IPRC8.5_stack, projPoly )
#predsProj <- raster::mask(predsProj, projPoly)
predsProjIPRC <- raster::stack(predsProjIPRC)
predsProjIPRC
```

```{r}
# projection over the globe under current conditions
myBiomodProj_IPRC <- BIOMOD_Projection(
bm.mod = all_models,
new.env = predsProjIPRC,
proj.name = 'future_IPRC',
selected.models = 'all',
binary.meth = 'TSS',
compress = 'xz',
clamping.mask = F,
output.format = '.img')
myBiomodProj_IPRC
```

```{r}
# make some plots sub-selected by str.grep argument
#plot(myBiomodProj_IPRC, str.grep = 'GLM')
```

```{r}
myBiomodProj_IPRC
```

```{r}
myBiomodEF_futureIPRC <- BIOMOD_EnsembleForecasting(
bm.em = myBiomodEM,
bm.proj = myBiomodProj_IPRC)
myBiomodEF_futureIPRC
```

```{r}
get_predictions(myBiomodEF_futureIPRC)
```

```{r}
## assign ensemble forecasting object with a new name for notation
## convenience
setwd(path.out.specific)
ensproj <- myBiomodEF_futureIPRC

## get the rasterstack of ensemble projections
ensproj.stack <- get_predictions(ensproj)

## define the output directory
outdir.ensproj <- paste0(spname, '_ensemble-projection_local_IPRC8.5')
dir.create(outdir.ensproj, showWarnings = FALSE, recursive = TRUE)

## save individual layers on the hard drive
purrr:::map(
names(ensproj.stack),
~ writeRaster(
ensproj.stack[[.x]],
overwrite = TRUE,
filename = file.path(outdir.ensproj, paste0(.x, '.tiff'))
)
)

list.files('outdir.ensproj')
```


```{r}
newdir <- paste(path.out.specific, "/", outdir.ensproj,  sep = "")
setwd(newdir)
final_raster_IPRC8.5 <- raster("Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo.tiff")

final_raster_IPRC8.5

plot(final_raster_IPRC8.5)
```
###save
```{r}
setwd(path.out.specific)
terra::writeRaster(final_raster_IPRC8.5, paste0(spname,"_IPRC8.5", "_local", "_final_ensemble_raster_local.img"),  overwrite = TRUE)
```

```{r}
maxValue(final_raster_NCAR8.5)
minValue(final_raster_NCAR8.5)
```



<!-- ###save -->
<!-- ```{r} -->
<!-- setwd(path.out.specific) -->
<!-- raster::writeRaster(combined, paste0(spname,"_bothfutures8.5", "_local", "_final_ensemble_raster23.img"),  overwrite = TRUE) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- names(combined) <- c(paste0(spname,"_bothfutures8.5", "_local", "_final_ensemble_raster.tif")) -->
<!-- x <- raster::writeRaster(combined, names(combined), overwrite=TRUE, filetype="GTiff") -->
<!-- ``` -->


```{r}
nowtime <- Sys.time()
elapsed = nowtime - starttime
print(elapsed)
```