---
title: "Andropogon_virginicus_Step2.2-globalmodeloutput"
output: html_document
date: "2023-09-03"
---

```{r}
starttime <- Sys.time()
```
# Name of Species being modelled
```{r}
spname <- "Andropogon_virginicus"
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,error=TRUE,message=FALSE)
```
```{r}
#get the needed packages
if(!require("pacman")){
	install.packages("pacman")
	library(pacman)}
p_load("dplyr", "PresenceAbsence", "DAAG", "ggplot2", 'tidyr', "terra", "sf", "knitr", "raster", "biomod2")
```
```{r}
# setting paths
  path.root <- "~/codingwork/fire"  # typical class root dir

  path.in.general <- paste(path.root, "/input", sep = "")
  
  path.in.specific <- paste(path.root, "/input/", spname, sep = "")
  path.out.specific <- paste(path.root, "/output/", spname, sep = "") 
  path.out.specific.ensemble.global <- paste(path.root, "/output/", spname, "/", spname, "_ensemble-projection_global", sep = "")
  path.out.specific.ensemble.global.current.HIdata <- paste(path.root, "/output/", spname, "/", spname, "_ensemble-projection_GLOBAL_current", sep = "")
  path.out.specific.ensemble.global.NCAR8.5 <- paste(path.root, "/output/", spname, "/", spname, "_ensemble-projection_GLOBAL_NCAR8.5", sep = "")
  path.out.specific.ensemble.global.IPRC8.5 <- paste(path.root, "/output/", spname, "/", spname, "_ensemble-projection_GLOBAL_IPRC8.5", sep = "")
  
  path.out.specific.ensemble.local <- paste(path.root, "/output/", spname, "/", spname, "_ensemble-projection_local", sep = "")
```
```{r}
# handy projection string
prj.wgs84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"  # epsg:4326
```

## Getting the points
```{r}
setwd(path.in.specific)
local.df <- read.csv(paste0(spname, "_tr_local_allvars.csv", sep = ""), header=T, sep=',', stringsAsFactors=F)
local.df # look at data - should include all variables
```

### Need to project these values as binary "present/not present" values

```{r}
pres.abs <- local.df[,1:4]
pres.abs
```

#COMPARING GLOBAL AND LOCAL MODELS AT CURRENT CLIMATE

## GLOBAL MODEL
## Getting the global model projected onto HI

```{r}
setwd(path.out.specific.ensemble.global.current.HIdata)
final_raster_global <- raster("Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo.tiff")

final_raster_global

#plot(final_raster_global)
```
Need to standardize the probabilities (from 0to 1)
```{r}
maxValue(final_raster_global)
minValue(final_raster_global)
```
```{r}

  layers <- {} # initialize (empty) list of raster layers
  for (i in 1:length(names(final_raster_global))) {
    m1 <- final_raster_global[[i]] # get a prob map
    m2 <- 1/maxValue(final_raster_global[[i]]) * final_raster_global[[i]] # standardize all probs to max=1
    m3 <- unlist(strsplit(names(final_raster_global[[i]]), "[.]")) # split prob layer name apart
    names(m2) <- paste(m3[1], ".", m3[2], sep = "")  # assign name to raster value
    assign(paste(m3[1], ".", m3[2], sep = ""), m2) # assign new name to standardized layer
    layers <- c(layers, get(paste(m3[1], ".", m3[2], sep = "")))
  }
  final_raster_global <- stack(layers)
  final_raster_global
```
```{r}
plot(final_raster_global)
maxValue(final_raster_global)
minValue(final_raster_global)
```
```{r}
t1 <- terra::extract(final_raster_global, pres.abs[c("decimalLongitude", "decimalLatitude")])
#t1[t1 == "NaN"] <- NA
head(t1, 10) # examine extracted matrix
```

### Finding Optim Threshold
```{r}
thresh_find_df <- cbind(pres.abs, t1) # bind to train dataframe
thresh_find_df # examine training data frame
```
```{r}
nrow(pres.abs)
```
```{r}
threshold_finder <- bm_FindOptimStat(metric.eval = "ROC", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo)
threshold_finder
```
```{r}
threshold_finder1 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .30)
threshold_finder2 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .35)
threshold_finder3 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .40)
threshold_finder4 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .45)
threshold_finder5 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .50)
threshold_finder6 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .55)
threshold_finder7 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .60)
threshold_finder8 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .65)
threshold_finder9 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .70)
threshold_finder10 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .75)

threshold_finderB <- rbind(threshold_finder1, threshold_finder2, threshold_finder3, threshold_finder4, threshold_finder5, threshold_finder6, threshold_finder7, threshold_finder8, threshold_finder9, threshold_finder10)
threshold_finderB
```
```{r}
threshold_finder1 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .450)
threshold_finder2 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .460)
threshold_finder3 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .470)
threshold_finder4 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .480)
threshold_finder5 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .490)
threshold_finder6 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .500)
threshold_finder7 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .510)
threshold_finder8 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .520)
threshold_finder9 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .530)
threshold_finder10 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .50)

threshold_finderC <- rbind(threshold_finder1, threshold_finder2, threshold_finder3, threshold_finder4, threshold_finder5, threshold_finder6, threshold_finder7, threshold_finder8, threshold_finder9, threshold_finder10)
threshold_finderC
```

### ASSIGN CUTOFF
```{r}
cutoff = 0.51


m <- c(0,cutoff,NA,
                 cutoff, 1.5, 1)
classmatrix <-  matrix(m, ncol=3, byrow=TRUE)
```
```{r}

final_raster_global_binary <- terra::classify(terra::rast(final_raster_global), classmatrix, include.lowest = TRUE)
plot(final_raster_global_binary)
```
```{r}
setwd(path.out.specific)
terra::writeRaster(final_raster_global_binary, paste0(spname,"_final_ensemble_raster_binary_currentHIdata.img"),  overwrite = TRUE)
```


## LOCAL MODEL
## Getting the LOCAL model projected onto HI

```{r}
setwd(path.out.specific.ensemble.local)
final_raster_local <- raster("Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo.tiff")

final_raster_local

#plot(final_raster_local)
```
Need to standardize the probabilities (from 0to 1)
```{r}
maxValue(final_raster_local)
minValue(final_raster_local)
```
```{r}

  layers <- {} # initialize (empty) list of raster layers
  for (i in 1:length(names(final_raster_local))) {
    m1 <- final_raster_local[[i]] # get a prob map
    m2 <- 1/maxValue(final_raster_local[[i]]) * final_raster_local[[i]] # standardize all probs to max=1
    m3 <- unlist(strsplit(names(final_raster_local[[i]]), "[.]")) # split prob layer name apart
    names(m2) <- paste(m3[1], ".", m3[2], sep = "")  # assign name to raster value
    assign(paste(m3[1], ".", m3[2], sep = ""), m2) # assign new name to standardized layer
    layers <- c(layers, get(paste(m3[1], ".", m3[2], sep = "")))
  }
  final_raster_local <- stack(layers)
  final_raster_local
```
```{r}
plot(final_raster_local)
maxValue(final_raster_local)
minValue(final_raster_local)
```
```{r}
t1 <- terra::extract(final_raster_local, pres.abs[c("decimalLongitude", "decimalLatitude")])
#t1[t1 == "NaN"] <- NA
head(t1, 10) # examine extracted matrix
```

### Finding Optim Threshold
```{r}
thresh_find_df <- cbind(pres.abs, t1) # bind to train dataframe
thresh_find_df # examine training data frame
```
```{r}
nrow(pres.abs)
```
```{r}
threshold_finder <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo)
threshold_finder
```
```{r}
threshold_finder1 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .30)
threshold_finder2 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .35)
threshold_finder3 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .40)
threshold_finder4 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .45)
threshold_finder5 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .50)
threshold_finder6 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .55)
threshold_finder7 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .60)
threshold_finder8 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .65)
threshold_finder9 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .70)
threshold_finder10 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .75)

threshold_finderB <- rbind(threshold_finder1, threshold_finder2, threshold_finder3, threshold_finder4, threshold_finder5, threshold_finder6, threshold_finder7, threshold_finder8, threshold_finder9, threshold_finder10)
threshold_finderB
```
```{r}
threshold_finder1 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .650)
threshold_finder2 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .660)
threshold_finder3 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .670)
threshold_finder4 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .680)
threshold_finder5 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .690)
threshold_finder6 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .700)
threshold_finder7 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .710)
threshold_finder8 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .620)
threshold_finder9 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .630)
threshold_finder10 <- bm_FindOptimStat(metric.eval = "TSS", thresh_find_df$scientificName, thresh_find_df$Andropogon.virginicus_EMmeanByROC_mergedData_mergedRun_mergedAlgo, threshold = .640)

threshold_finderC <- rbind(threshold_finder1, threshold_finder2, threshold_finder3, threshold_finder4, threshold_finder5, threshold_finder6, threshold_finder7, threshold_finder8, threshold_finder9, threshold_finder10)
threshold_finderC
```

### ASSIGN CUTOFF
```{r}
cutoff = 0.5


m <- c(0,cutoff,NA,
                 cutoff, 1.5, 1)
classmatrix <-  matrix(m, ncol=3, byrow=TRUE)


final_raster_local_binary <- terra::classify(terra::rast(final_raster_local), classmatrix, include.lowest = TRUE)
plot(final_raster_global_binary)
plot(final_raster_local_binary)

```
```{r}
setwd(path.out.specific)
terra::writeRaster(final_raster_local_binary, paste0(spname,"_final_ensemble_raster_binary.img"),  overwrite = TRUE)
```

## COMPARE

```{r}

```







































taking the average of NCAR and IPRC
```{r}
combined <- raster::stack(final_raster_NCAR8.5, final_raster_IPRC8.5)
combined <- mean(combined)
combined
```

```{r}

  layers <- {} # initialize (empty) list of raster layers
  for (i in 1:length(names(combined))) {
    m1 <- combined[[i]] # get a prob map
    m2 <- 1/maxValue(combined[[i]]) * combined[[i]] # standardize all probs to max=1
    m3 <- unlist(strsplit(names(combined[[i]]), "[.]")) # split prob layer name apart
    names(m2) <- paste(m3[1], ".", m3[2], sep = "")  # assign name to raster value
    assign(paste(m3[1], ".", m3[2], sep = ""), m2) # assign new name to standardized layer
    layers <- c(layers, get(paste(m3[1], ".", m3[2], sep = "")))
  }
  combined <- stack(layers)
  combined
```
```{r}
plot(combined)
```
























```{r}
nowtime <- Sys.time()
elapsed = nowtime - starttime
print(elapsed)
```
