---
title: "Cenchrus setaceus - Global Model Outputs"
author: "Kelsey Brock"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---
# Name of Species being modelled
```{r}
spname <- "Cenchrus_setaceus"
```
```{r}
futureclim <- "RCP8.5"
fclim <- 85
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,error=TRUE,message=FALSE)
```
```{r}
#get the needed packages
if(!require("pacman")){
	install.packages("pacman")
	library(pacman)}
p_load("dplyr", "PresenceAbsence", "DAAG", "ggplot2", 'tidyr', "terra", "sf", "knitr", "dismo", "gam", "randomForest", "gbm", "biomod2")
```
```{r}
# setting paths
  path.root <- "~/codingwork/fire"  # typical class root dir
  path.in.general <- paste(path.root, "/input", sep = "")
  path.in.specific <- paste(path.root, "/input/", spname, sep = "")
  path.out.specific <- paste(path.root, "/output/", spname, sep = "") 
```
```{r}
# handy projection string
prj.wgs84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"  # epsg:4326
```

```{r}
#memory.limit(35000)
```


####  Get the Occurrence Data

```{r}
setwd(path.in.specific)
global.df <- read.csv(paste(spname, "_tr_global_allvars.csv", sep = ""), header=T, sep=',', stringsAsFactors=F)
head(global.df) # look at data - should include all variables
```


```{r}
table(global.df$Species) # examine  frequencies
```

### Get selected variables during data exploration phase
```{r}
setwd(path.in.specific)
keep <- get(load(file = paste0( spname, "_list_picked_vars_global.RData")))
keep
```

### Import raster of current and future climate scenarios

#### Current


```{r}
setwd(path.in.specific)
#rasterOptions(maxmemory=1e+07)
#stack <- get(load("bioclimstack.RData"))
stack <- terra::rast( paste0(spname, "_bioclimstack.tif"))
stack
```


```{r}
stack <- terra::subset(stack, keep)
#stack3 <- stack(stack)
stack

#writeRaster(stack3,"tempstack.img", format = "HFA", overwrite = TRUE)
```


#### Future RCP 8.5

Un-comment this for the first use to get future climate data


<!-- ```{r} -->
<!-- setwd(path.in.general) -->
<!-- envs<-terra::rast("wc2.1.30s_bioc_ACCESS-ESM1-5_ssp585_2081-2100.tif") -->
<!-- envs -->
<!-- ``` -->


<!-- ```{r} -->
<!-- setwd(path.in.general) -->
<!-- elev <- terra::rast("wc2.1_30s_elev.tif") -->
<!-- elev -->
<!-- ``` -->
<!-- ```{r} -->
<!-- combined <- c(envs, elev) -->

<!-- names(combined) <- c("bio1",  "bio2",  "bio3",  "bio4",  "bio5",  "bio6",  "bio7",  "bio8",  "bio9", "bio10", "bio11", "bio12", "bio13", "bio14", "bio15", "bio16", "bio17", "bio18", "bio19", "alt"  ) -->
<!-- names(combined) -->

<!-- combined -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #global.bufbbox <- extent(min(global.pres$decimalLongitude) - (0.00833 * buffer), max(global.pres$decimalLongitude) + (0.00833 * buffer), -->
<!-- #               min(global.pres$decimalLatitude) - (0.00833 * buffer), max(global.pres$decimalLatitude) + (0.00833 * buffer)) -->

<!-- #making a bbox of the entire world, excluding the arctic and antarctica -->

<!-- xmin = -179.9999 -->
<!-- xmax =  179.9999 -->
<!-- ymin = -61 -->
<!-- ymax =  61 -->

<!-- global.bufbbox <- extent(xmin, xmax, ymin, ymax) -->

<!-- global.bufbbox -->
<!-- ``` -->

<!-- ```{r} -->
<!-- combined <- terra::crop(combined, global.bufbbox) -->
<!-- combined -->
<!-- #raster::crop(raster::rasterize(global.bufbboxSP, template1), global.bufbboxSP) -->

<!-- ``` -->


<!-- ```{r} -->
<!-- setwd(path.in.general) -->
<!-- #save(combined, file = "cmip_eightfive.RData") # save .RData -->
<!-- writeRaster(combined, filename ="cmip_eightfive.tif", overwrite = T) # save as .img file -->
<!-- ``` -->

```{r}
setwd(path.in.general)
cmip_eightfive <- terra::rast("cmip_eightfive.tif")

cmip_eightfive <- terra::subset(cmip_eightfive, keep)
cmip_eightfive
```



<!-- ## clipping to the buffered pts too -->
<!-- ```{r} -->
<!-- setwd(path.in.specific) -->
<!-- global.bufptSF <- get(load(file = paste0(spname, "_global.bufptSF.RData"))) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- cmip_eightfive_buffered <- terra::crop(cmip_eightfive, global.bufptSF) -->
<!-- cmip_eightfive_buffered -->
<!-- #raster::crop(raster::rasterize(global.bufbboxSP, template1), global.bufbboxSP) -->

<!-- ``` -->


# BIOMOD2 detour


```{r}
global.df
```

```{r}
responsevar <- c("FNETID", "Species", "tr.Lon", "tr.Lat" )
responsevar
keep
```

```{r}
global.df <- global.df %>% dplyr::select(responsevar, keep) 
global.df
```


Getting a weird error!?!?  this is probably the chunk that's the problem.  Double check indexing
```{r}
# Clean data
myrespvar <- as.numeric(global.df$Species)
myRespName <- spname
myexplvar <- global.df[, 5:length(global.df)]
myresp.xy <- global.df[, 3:4]
clean_data <- BIOMOD_FormatingData(resp.var=myrespvar, expl.var=myexplvar, resp.xy=myresp.xy, resp.name=myRespName)
```

```{r}
clean_data
```

```{r}
my_models <- c("GLM","GAM","MAXENT.Phillips","RF","GBM") # Algorithms to fit the models
```

```{r}
# Settings for the model algorithms
model_opt <- BIOMOD_ModelingOptions(GLM=list(type='quadratic', interaction.level=0),GBM=list(n.trees=1000),MAXENT.Phillips = list(path_to_maxent.jar = "C:/RFolder/R/win-library/4.1/dismo/java/maxent.jar",product=FALSE),GAM=list(k=3))

```

```{r}
all_models <- BIOMOD_Modeling(data=clean_data, models=my_models, models.options=model_opt, NbRunEval = 6, DataSplit = 80, VarImport = 3, do.full.models = F, modeling.id = "global")
```

```{r}
all_models
```

```{r}
# get all models evaluation
myBiomodModelEval <- get_evaluations(all_models)
# print the dimnames of this object
dimnames(myBiomodModelEval)
```
```{r}
# let's print the ROC scores of all selected models
myBiomodModelEval["TSS","Testing.data",,,]
```

```{r}
# let's print the ROC scores of all selected models
myBiomodModelEval["ROC","Testing.data",,,]
```
```{r}
# print variable importances
get_variables_importance(all_models)
```


<!-- ```{r} -->
<!-- df <- local.df %>% dplyr::select(!(FNETID:tr.Lat)) %>% dplyr::select(!(yweights))# %>% dplyr::rename(x = tr.Lon, y = tr.Lat) -->
<!-- #colnames(df) -->
<!-- ``` -->




```{r}
all_models
```


### Project to HI



```{r}
xmin = -160.5613
xmax =  -154.7061
ymin = 17.9578 
ymax =  22.3496

projPoly <- terra::ext(xmin, xmax, ymin, ymax)

projPoly
```

```{r}
predsProj <- terra::crop(stack, projPoly)
#predsProj <- raster::mask(predsProj, projPoly)
#predsProj <- raster::stack(predsProj)
predsProj
```
```{r}
library(raster)
predsProj <- raster::stack(predsProj)
  predsProj
```


```{r}
# projection over the globe under current conditions
myBiomodProj <- BIOMOD_Projection(
modeling.output = all_models,
new.env = predsProj,
proj.name = 'current',
selected.models = 'all',
binary.meth = 'TSS',
compress = 'xz',
clamping.mask = F,
output.format = '.img')
myBiomodProj
```

```{r}
# make some plots sub-selected by str.grep argument
plot(myBiomodProj, str.grep = 'GLM')
```
```{r}
# make some plots sub-selected by str.grep argument
plot(myBiomodProj, str.grep = 'GAM')
```
```{r}
# make some plots sub-selected by str.grep argument
plot(myBiomodProj, str.grep = 'MAXENT.Phillips')
```
```{r}
# make some plots sub-selected by str.grep argument
plot(myBiomodProj, str.grep = 'RF')
```
```{r}
# make some plots sub-selected by str.grep argument
plot(myBiomodProj, str.grep = 'GBM')
```

```{r}
myBiomodProj
```

```{r}
myBiomodEM <- BIOMOD_EnsembleModeling(
modeling.output = all_models,
chosen.models = 'all',
em.by='all',
eval.metric = c('ROC'),
eval.metric.quality.threshold = c(0.63),
prob.mean = T,
prob.cv = T,
prob.ci = T,
prob.ci.alpha = 0.03,
prob.median = T,
committee.averaging = T,
prob.mean.weight = T,
prob.mean.weight.decay = 'proportional' )
```
```{r}
# get evaluation scores
get_evaluations(myBiomodEM)
```


```{r}
myBiomodEF_current <- BIOMOD_EnsembleForecasting(
EM.output = myBiomodEM,
projection.output = myBiomodProj)
myBiomodEF_current
```

```{r}
plot(myBiomodEF_current, str.grep = 'meanBy')
```
```{r}
get_predictions(myBiomodEF_current)
```


```{r}
## assign ensemble forecasting object with a new name for notation
## convenience
setwd(path.out.specific)
ensproj <- myBiomodEF_current

## get the rasterstack of ensemble projections
ensproj.stack <- get_predictions(ensproj)

## define the output directory
outdir.ensproj <- paste0(spname, '_ensemble-projection')
dir.create(outdir.ensproj, showWarnings = FALSE, recursive = TRUE)

## save individual layers on the hard drive
purrr:::map(
names(ensproj.stack),
~ writeRaster(
ensproj.stack[[.x]],
overwrite = TRUE,
filename = file.path(outdir.ensproj, paste0(.x, '.tiff'))
)
)

list.files('outdir.ensproj')
```

```{r}
newdir <- paste(path.out.specific, "/", outdir.ensproj,  sep = "")
setwd(newdir)
final_raster <- raster("Cenchrus.setaceus_EMmeanByROC_mergedAlgo_mergedRun_mergedData.tiff")
```

```{r}
plot(final_raster)
```

```{r}
setwd(path.out.specific)
writeRaster(final_raster, paste0(spname,"_final_ensemble_raster.img"), format = "HFA", overwrite = TRUE)
```


## FUTURE CLIMATE

```{r}
cmip3 <- terra(cmip_eightfive)
cmip3
```

```{r}
predsProj_cmip <- raster::crop(cmip3, projPoly)
#predsProj <- raster::mask(predsProj, projPoly)
predsProj_cmip <- raster::stack(predsProj_cmip)
predsProj_cmip
```

```{r}
# projection over the globe under current conditions
myBiomodProj <- BIOMOD_Projection(
modeling.output = all_models,
new.env = cmip3,
proj.name = 'current',
selected.models = 'all',
binary.meth = 'TSS',
compress = 'xz',
clamping.mask = F,
output.format = '.img')
myBiomodProj
```

```{r}
# make some plots sub-selected by str.grep argument
plot(myBiomodProj) #, str.grep = 'GBM')
```
```{r}
myBiomodProj
```

```{r}
myBiomodEM <- BIOMOD_EnsembleModeling(
modeling.output = all_models,
chosen.models = 'all',
em.by='all',
eval.metric = c('TSS'),
#eval.metric.quality.threshold = c(0.4),
prob.mean = T,
prob.cv = T,
prob.ci = T,
prob.ci.alpha = 0.03,
prob.median = T,
committee.averaging = T,
prob.mean.weight = T,
prob.mean.weight.decay = 'proportional' )
```
```{r}
# get evaluation scores
get_evaluations(myBiomodEM)
```

```{r}
myBiomodEF_future <- BIOMOD_EnsembleForecasting(
EM.output = myBiomodEM,
projection.output = myBiomodProj)
myBiomodEF_future
```

```{r}
plot(myBiomodEF_current, str.grep = 'Cenchrus.setaceus_EMmeanByTSS_mergedAlgo_mergedRun_mergedData')
```
```{r}
get_predictions(myBiomodEF_current)
```


```{r}
## assign ensemble forecasting object with a new name for notation
## convenience
setwd(path.out.specific)
ensproj <- myBiomodEF_future

## get the rasterstack of ensemble projections
ensproj.stack <- get_predictions(ensproj)

## define the output directory
outdir.ensproj <- paste0(spname, futureclim, '_ensemble-projection')
dir.create(outdir.ensproj, showWarnings = FALSE, recursive = TRUE)

## save individual layers on the hard drive
purrr:::map(
names(ensproj.stack),
~ writeRaster(
ensproj.stack[[.x]],
overwrite = TRUE,
filename = file.path(outdir.ensproj, paste0(.x, '.tiff'))
)
)

list.files('outdir.ensproj')
```

```{r}
newdir <- paste(path.out.specific, "/", outdir.ensproj,  sep = "")
setwd(newdir)
final_raster <- raster("Cenchrus.setaceus_EMmeanByTSS_mergedAlgo_mergedRun_mergedData.tiff")
```

```{r}
plot(final_raster)
```

```{r}
setwd(path.out.specific)
writeRaster(final_raster, paste0(spname, futureclim, "_final_ensemble_raster.img"), format = "HFA", overwrite = TRUE)
```

