---
title: "- local Model Outputs"
author: "Kelsey Brock"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

# Name of Species being modelled
```{r}
# provide species name in format "Genus_species"
spname <- ""
```
```{r}
# provide relative concentration pathway for climate change data
futureclim <- "RCP8.5"
fclim <- 85
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,error=TRUE,message=FALSE)
```
```{r}
#get the needed packages
if(!require("pacman")){
	install.packages("pacman")
	library(pacman)}
p_load("dplyr", "PresenceAbsence", "DAAG", "ggplot2", 'tidyr', "raster", "sf", "knitr", "dismo", "gam", "randomForest", "gbm", "biomod2")
```
```{r}
# setting paths
  path.root <- "~/codingwork/fire"  # typical class root dir
  path.in.general <- paste(path.root, "/input", sep = "")
  path.in.specific <- paste(path.root, "/input/", spname, sep = "")
  path.out.specific <- paste(path.root, "/output/", spname, sep = "") 
```
```{r}
# handy projection string
prj.wgs84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"  # epsg:4326
```
```{r}
memory.limit(size=25000)
```

####  Get the Occurrence Data

```{r}
setwd(path.in.specific)
local.df <- read.csv(paste(spname, "_tr_local_allvars.csv", sep = ""), header=T, sep=',', stringsAsFactors=F)
local.df # look at data - should include all variables
```
```{r}
table(local.df$Species) # examine  frequencies
```
### Get Global Projection for Pseudoabsence down-weighting

```{r}
setwd(path.out.specific)
GlobalProj <- get(load(file = paste0( spname, "_probSTD.mean.ens.RData")))
GlobalProj
```

### Assigning weight to pseudo absences according to: 
Gallien L, Douzet R, Pratte S, Zimmermann NE, Thuiller W (2012) Invasive species distribution models – how violating the equilibrium assumption can create new insights. Global Ecology and Biogeography 21: 1126–1136. https://doi.org/10.1111/j.1466-8238.2012.00768.x 

```{r}
weights_gl <- 1/(1+((GlobalProj/(GlobalProj - 1))^2)) # inverse logistic transformation

xmin = -160.3722
xmax =  -154.5714
ymin =  18.7084
ymax =  22.4615

local.bufbbox <- extent(xmin, xmax, ymin, ymax)

weights_gl <- crop(weights_gl, local.bufbbox )

weights_gl
```

```{r}
pres <- subset(local.df, Species == 1)
pres$yweights <- 1

abs <- subset(local.df, Species == 0)
abs
```

```{r}
weight_values_ab <- raster::extract(weights_gl, cbind(abs$tr.Lon, abs$tr.Lat))

abs$yweights <- weight_values_ab
```
```{r}
local.df <- rbind(pres, abs)
local.df
```

### Get  variables selected during data exploration phase
```{r}
setwd(path.in.specific)
keep <- get(load(file = paste0( spname, "_list_picked_vars_local.RData")))
keep
```

```{r}
header <- c("FNETID", "Species", "tr.Lon", "tr.Lat")
oo <- c(header, keep)
picked.df <- local.df %>% dplyr::select(oo)
head(picked.df, 2)
```

```{r}
pres <- subset(picked.df, Species == 1)
pres$yweights <- 1

abs <- subset(picked.df, Species == 0)
```

```{r}
weight_values_ab <- raster::extract(weights_gl, cbind(abs$tr.Lon, abs$tr.Lat))

abs$yweights <- weight_values_ab
```
```{r}
picked.df <- rbind(pres, abs)
picked.df
```

### Import raster of current and future climate scenarios

#### Current

```{r}
setwd(path.in.general)
stack <- get(load("HI_bioclimstack.RData"))

xmin = -160.3722
xmax =  -154.5714
ymin =  18.7084
ymax =  22.4615

local.bufbbox <- extent(xmin, xmax, ymin, ymax)


stack <- crop(stack, local.bufbbox )
stack

```
```{r}
path.in.layers <- paste(path.in.general, "/yearly_and_seasonal_projections_HI/",sep = "") 
path.in.layers

```

```{r}
setwd(path.in.layers)
# Extract all .rds files
list <- list.files( pattern =".tif", full.names = TRUE)

# Define pattern for grepl
pattern <- paste(keep, sep="", collapse="|")


# Now we can subset list with the following
pickedvars <- list[grepl(pattern,list)]

stack1 <- raster::stack(pickedvars)
stack1
```



#### Future RCP 8.5

```{r}
#cmip_eightfive <- raster::getData(name = 'CMIP5', var = 'bio', res = 2.5,
#        rcp = fclim, model = 'AC', year = 70,
#        path = paste0(path.in.general, "/bioclims_CMIP5_ssp585"))
```

```{r}
# get WorldClim bioclimatic variable rasters
#elev <- raster::getData(name = "worldclim", var = "alt", res = 2.5, lat = , lon = )
#elev
```

Using Hi-specific climate variables, and following bioclim namings out of convenience

```{r}
path.preds <- paste(path.root, "/input/yearly_and_seasonal_projections_HI",  sep = "")
setwd(path.preds)
future <- list.files(pattern = "rcp85.tif") # list of .img files; $ strips extra
cmip_eightfive <- stack(future)# examine
cmip_eightfive
```

```{r}
#cmip_eightfive <- raster::stack(cmip_eightfive, elev)
#names(cmip_eightfive)
```

```{r}
names(cmip_eightfive) <- names(stack)

xmin = -160.3722
xmax =  -154.5714
ymin =  18.7084
ymax =  22.4615

local.bufbbox <- extent(xmin, xmax, ymin, ymax)


cmip_eightfive <- crop(cmip_eightfive, local.bufbbox )

cmip_eightfive
```

# BIOMOD2 

```{r}
my_models <- c("GLM","GAM","MAXENT.Phillips","RF","GBM") # Algorithms to fit the models
```

```{r}
# Settings for the model algorithms
model_opt <- BIOMOD_ModelingOptions(GLM=list(type='quadratic', interaction.level=0),GBM=list(n.trees=1000),MAXENT.Phillips = list(path_to_maxent.jar = "C:/Users/kcbro/Documents/R/win-library/4.0/dismo/java/maxent.jar",product=FALSE),GAM=list(k=3))

```

```{r}
local.df <- dplyr::select(local.df, !(isle.y))
local.df <- dplyr::select(local.df, !(isle.x))
local.df <- na.omit(local.df)
local.df
```

PICKED VARS ONLY

```{r}
local.df <- na.omit(picked.df) # switch to local.df if you want to experiment with all vars.
local.df
```

```{r}
# Clean data
myrespvar <- as.numeric(local.df$Species)
myRespName <- spname
#myexplvar <- local.df[, 7:18]
#myresp.xy <- local.df[, 5:6]

## FOR PICKED!! Be careful here
myexplvar <- local.df[, 5:11]
myresp.xy <- local.df[, 3:4]

clean_data <- BIOMOD_FormatingData(resp.var=myrespvar, expl.var=myexplvar, resp.xy=myresp.xy, resp.name=myRespName)
```
```{r}
clean_data
```

```{r}
all_models <- BIOMOD_Modeling(data=clean_data, models=my_models, models.options=model_opt, NbRunEval = 4, DataSplit = 80, VarImport = 3, do.full.models = F, modeling.id = "local", Yweights=local.df$yweights)

```
```{r}
all_models
```

```{r}
# get all models evaluation
myBiomodModelEval <- get_evaluations(all_models)
# print the dimnames of this object
dimnames(myBiomodModelEval)
```
```{r}
# let's print the TSS scores of Random Forest
myBiomodModelEval["TSS","Testing.data","GLM",,]
myBiomodModelEval["TSS","Testing.data","GAM",,]
myBiomodModelEval["TSS","Testing.data","MAXENT.Phillips",,]
myBiomodModelEval["TSS","Testing.data","RF",,]
myBiomodModelEval["TSS","Testing.data","GBM",,]
```

```{r}
# let's print the ROC scores of all selected models
myBiomodModelEval["ROC","Testing.data",,,]
```

```{r}
# print variable importances
get_variables_importance(all_models)
```


```{r}
stack3 <- stack(stack1) #stack
stack3
```

```{r}
all_models
```

```{r}
# projection over the globe under current conditions
myBiomodProj <- BIOMOD_Projection(
modeling.output = all_models,
new.env = stack3,
proj.name = 'current',
selected.models = 'all',
binary.meth = 'TSS',
compress = 'xz',
clamping.mask = F,
output.format = '.img')
myBiomodProj
```
```{r}
stack3
```

```{r}
# make some plots sub-selected by str.grep argument
plot(myBiomodProj) #, str.grep = 'GBM')
```
```{r}
myBiomodProj
```

```{r}
myBiomodEM <- BIOMOD_EnsembleModeling(
modeling.output = all_models,
chosen.models = 'all',
em.by='all',
eval.metric = c('TSS'),
#eval.metric.quality.threshold = c(0.4),
prob.mean = T,
prob.cv = T,
prob.ci = T,
prob.ci.alpha = 0.03,
prob.median = T,
committee.averaging = T,
prob.mean.weight = T,
prob.mean.weight.decay = 'proportional' )
```
```{r}
# get evaluation scores
get_evaluations(myBiomodEM)
```

```{r}
myBiomodEF_current <- BIOMOD_EnsembleForecasting(
EM.output = myBiomodEM,
projection.output = myBiomodProj)
myBiomodEF_current
```

```{r}
plot(myBiomodEF_current, str.grep = 'Species.name_EMmeanByTSS_mergedAlgo_mergedRun_mergedData')
```
```{r}
get_predictions(myBiomodEF_current)
```


```{r}
## assign ensemble forecasting object with a new name for notation
## convenience
setwd(path.out.specific)
ensproj <- myBiomodEF_current

## get the rasterstack of ensemble projections
ensproj.stack <- get_predictions(ensproj)

## define the output directory
outdir.ensproj <- paste0(spname, '_ensemble-projection')
dir.create(outdir.ensproj, showWarnings = FALSE, recursive = TRUE)

## save individual layers on the hard drive
purrr:::map(
names(ensproj.stack),
~ writeRaster(
ensproj.stack[[.x]],
overwrite = TRUE,
filename = file.path(outdir.ensproj, paste0(.x, '.tiff'))
)
)

list.files('outdir.ensproj')
```

```{r}
newdir <- paste(path.out.specific, "/", outdir.ensproj,  sep = "")
setwd(newdir)
final_raster <- raster("Species.name_EMmeanByTSS_mergedAlgo_mergedRun_mergedData.tiff")
```

```{r}
plot(final_raster)
```

```{r}
setwd(path.out.specific)
writeRaster(final_raster, paste0(spname,"_final_ensemble_raster.img"), format = "HFA", overwrite = TRUE)
```


## FUTURE CLIMATE

```{r}
cmip3 <- stack(cmip_eightfive)
cmip3
```
```{r}
cmip3 <- raster::subset(cmip3, colnames(local.df[5:11]))
cmip3
```


```{r}
# projection over the globe under current conditions
myBiomodProj <- BIOMOD_Projection(
modeling.output = all_models,
new.env = cmip3,
proj.name = 'current',
selected.models = 'all',
binary.meth = 'TSS',
compress = 'xz',
clamping.mask = F,
output.format = '.img')
myBiomodProj
```


```{r}
# make some plots sub-selected by str.grep argument
plot(myBiomodProj) #, str.grep = 'GBM')
```
```{r}
myBiomodProj
```

```{r}
myBiomodEM <- BIOMOD_EnsembleModeling(
modeling.output = all_models,
chosen.models = 'all',
em.by='all',
eval.metric = c('TSS'),
eval.metric.quality.threshold = c(0.65),
prob.mean = T,
prob.cv = T,
prob.ci = T,
prob.ci.alpha = 0.03,
prob.median = T,
committee.averaging = T,
prob.mean.weight = T,
prob.mean.weight.decay = 'proportional' )
```
```{r}
# get evaluation scores
get_evaluations(myBiomodEM)
```

```{r}
myBiomodEF_future <- BIOMOD_EnsembleForecasting(
EM.output = myBiomodEM,
projection.output = myBiomodProj)
myBiomodEF_future
```

```{r}
plot(myBiomodEF_current, str.grep = 'Species.name_EMmeanByTSS_mergedAlgo_mergedRun_mergedData')
```
```{r}
get_predictions(myBiomodEF_current)
```


```{r}
## assign ensemble forecasting object with a new name for notation
## convenience
setwd(path.out.specific)
ensproj <- myBiomodEF_future

## get the rasterstack of ensemble projections
ensproj.stack <- get_predictions(ensproj)

## define the output directory
outdir.ensproj <- paste0(spname, futureclim, '_ensemble-projection')
dir.create(outdir.ensproj, showWarnings = FALSE, recursive = TRUE)

## save individual layers on the hard drive
purrr:::map(
names(ensproj.stack),
~ writeRaster(
ensproj.stack[[.x]],
overwrite = TRUE,
filename = file.path(outdir.ensproj, paste0(.x, '.tiff'))
)
)

list.files('outdir.ensproj')
```

```{r}
newdir <- paste(path.out.specific, "/", outdir.ensproj,  sep = "")
setwd(newdir)
final_raster <- raster("Species.name_EMmeanByTSS_mergedAlgo_mergedRun_mergedData.tiff")
```

```{r}
plot(final_raster)
```

```{r}
setwd(path.out.specific)
writeRaster(final_raster, paste0(spname, futureclim, "_final_ensemble_raster.img"), format = "HFA", overwrite = TRUE)
```





