---
title: " - Preparing Global Model Inputs"
author: "Kelsey Brock"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---
# Name of Species being modelled
```{r}
# place species name in format: "Genus_species"
spname <- ""
```
```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,error=TRUE,message=FALSE)
```
```{r}
#get the needed packages
if(!require("pacman")){
	install.packages("pacman")
	library(pacman)}
p_load("dplyr", "tidyr", "ggplot2", "raster",  "sf", "sp", "maptools", "CoordinateCleaner",  "spThin", "modeest", "beepr")
```
```{r}
# setting paths
  path.root <- "~/codingwork/fire"  # typical class root dir
  path.in.general <- paste(path.root, "/input", sep = "")
  path.in.specific <- paste(path.root, "/input/", spname, sep = "")
  path.out <- paste(path.root, "/output/", spname, sep = "") 
```
```{r}
# handy projection string
prj.wgs84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"  # epsg:4326
```
```{r}
memory.limit(25000)
```


####  Get the Occurrence Data
```{r}
pres <- read.csv(paste(path.in.general, "/occurrence_data/", spname, ".csv", sep = ""), header=T, sep=',', stringsAsFactors=F)
head(pres) # look at data
```
```{r}
# select only herbarium data if you are unconfident in other occurrence info
pres <- subset(pres, pres$basisOfRecord == "PRESERVED_SPECIMEN")
pres$Species <- 1
pres
```

e(pres$Species)
```
How many 1s (presences)  do we have?
```{r}
nrow(pres)
```


### Clean Coordinates

```{r}
flagged <- clean_coordinates(x = pres, species = "Species", lon = "decimalLongitude", lat = "decimalLatitude", 
                             tests = c( "centroids","equal", "seas", "zeros"))
summary(flagged)
```
```{r}
table(flagged$.summary)
```
```{r}
pres <- subset(flagged, flagged$.summary == TRUE)
dim(pres)
```
#### Spatial thinning (takes a long time)

```{r}
output <- spThin::thin(pres, lat.col = 'decimalLatitude', long.col = 'decimalLongitude', spec.col = "Species", thin.par = 50, reps = 3, locs.thinned.list.return = TRUE, write.files = FALSE, verbose = TRUE)
```

Since spThin did 100 iterations, there are 100 different variations of
how it thinned the occurrence localities. As there is a stochastic
element in the algorithm, some iterations may include more localities
than the others, and we need to make sure we maximize the number of
localities we proceed with.

```{r}
# find the iteration that returns the max number of occurrences
maxThin <- which(sapply(output, nrow) == max(sapply(output, nrow)))
# if there's more than one max, pick the first one
maxThin <- output[[ifelse(length(maxThin) > 1, maxThin[1], maxThin)]]  
# subset occs to match only thinned occs
pres <- pres[as.numeric(rownames(maxThin)),]  
pres
```
save the raw points that were used
```{r}
setwd(path.in.specific)
write.csv(pres, paste0(spname, "_occurrences_used.csv"), row.names = FALSE)
```

```{r}
setwd(path.in.specific)
pres <- read.csv(paste(spname, "_occurrences_used.csv", sep = ""), header=T, sep=',', stringsAsFactors=F)
```
```{r}
nrow(pres)
pres <- pres %>% dplyr::select(Species, decimalLongitude, decimalLatitude)
head(pres)
```

#### Converting to a spatial object with a geometry field
```{r}
pres <- st_as_sf(pres, coords = c("decimalLongitude", "decimalLatitude"), crs = prj.wgs84, remove = FALSE)
class(pres)
head(pres, 4)
```

#### Subset into local versus Global dataset

```{r}
setwd(path.in.general)
#This polygon is only for the main hawaiian islands only
islandpolys <- st_read("Coastline.shp")
```
```{r}
local.pres <- st_join(pres, islandpolys, join = st_intersects, left = FALSE ) %>% dplyr::select(Species, decimalLongitude, decimalLatitude, isle,	geometry)
unique(local.pres$isle)
dim(local.pres)
head(local.pres)
```
#### Check:

```{r}
plot(st_geometry(islandpolys), border = "darkgrey", axes = F)
plot(st_geometry(local.pres),  add= T, col = "red", pch = 16)
```
```{r}
worldmap <- borders("world", colour = "white", fill = "gray75")
map_world <- ggplot() +
                worldmap +
                geom_point(local.pres, mapping = aes(x = decimalLongitude, y = decimalLatitude), col = "red")
map_world
```
```{r}
global.pres <- st_join(pres, islandpolys, join = st_intersects,  left = TRUE ) %>% 
  subset(is.na(isle)) %>% 
  dplyr::select(Species, decimalLongitude, decimalLatitude, isle,	geometry)
unique(global.pres$isle)
dim(global.pres)
head(global.pres)
```
```{r}
plot(st_geometry(islandpolys), border = "darkgrey", axes = F)
plot(st_geometry(global.pres),  add= T, col = "red", pch = 16)
```
```{r}
worldmap <- borders("world", colour = "white", fill = "gray75")
map_world <- ggplot() +
                worldmap +
                geom_point(global.pres, mapping = aes(x = decimalLongitude, y = decimalLatitude), col = "red")
map_world
```
# GLOBAL MODEL

#### making a bounding box (with  buffered points)

```{r}
# buffer around the points within which to ultimately choose pseudoabsences in m
buffer = 5000000
```
```{r}
#global.bufbbox <- extent(min(global.pres$decimalLongitude) - (0.00833 * buffer), max(global.pres$decimalLongitude) + (0.00833 * buffer),
#               min(global.pres$decimalLatitude) - (0.00833 * buffer), max(global.pres$decimalLatitude) + (0.00833 * buffer))

#making a bbox of the entire world, excluding the arctic and antarctica

xmin = -179.9999 
xmax =  179.9999
ymin = -66.9999 
ymax =  66.9999

global.bufbbox <- extent(xmin, xmax, ymin, ymax)

global.bufbbox
```
```{r}
global.bufbboxSF <- st_as_sfc(st_bbox(global.bufbbox, crs = prj.wgs84))
global.bufbboxSF
```
```{r}
plot(st_geometry(global.bufbboxSF), axes = T)
plot(st_geometry(global.pres), add= T, col = "darkgreen")
```

#### Making a SF polygon of the buffered points

```{r}
#buffering each point by 50km (remember, each km is approx .0083 degrees)
global.bufptSF <- sf::st_buffer(global.pres, dist = (0.00833 * buffer))

plot(st_geometry(global.bufptSF), axes = T)
plot(st_geometry(global.bufptSF), add= T, col = "darkgreen")
```


```{r}
#union-ing the points into a single polygon
global.bufptSF <- sf::st_union(global.bufptSF)
plot(st_geometry(global.bufptSF), axes = T)
```


#### Saving
(This will be domain where the statistical model will be projected later on)

```{r}
setwd(path.in.specific)
save("global.bufptSF", file = paste0(spname, "_global.bufptSF.RData"))
save("global.bufbboxSF", file = paste0(spname, "_global.bufbbox.RData"))
```


## Building a Global "Fishnet"

This creates a dataframe of gridcells to associate with presence points, so that each presence point can be assigned a unique FNET ID

#### converting buffered pt polygon to raster using a template raster file
```{r}
setwd(path.in.general)
template1 <- raster("gpw-v4-land-water-area-rev11_landareakm_2pt5_min_tif/gpw_v4_land_water_area_rev11_landareakm_2pt5_min.tif")
template1
```

Let’s make sure the projections are the same:
```{r}
st_crs(template1)$proj4string == st_crs(global.bufbboxSF)$proj4string
```
```{r}
st_crs(global.bufbboxSF)$proj4string == st_crs(global.bufptSF)$proj4string
```
```{r}
global.bufbboxSP <- as_Spatial(global.bufbboxSF) # converts to spatial polygon class
global.bufboxR <- raster::crop(raster::rasterize(global.bufbboxSP, template1), global.bufbboxSP) # cropping to bbox
global.bufboxR
```
```{r}
plot(global.bufboxR, legend = T)
```
#### Save
```{r}
setwd(path.in.general)
writeRaster(global.bufboxR,"global.bboxR.img", format = "HFA", overwrite = TRUE)
```
```{r}
global.bufptSP <- as_Spatial(global.bufptSF) # converts to spatial polygon class
global.bufptR <- crop(rasterize(global.bufptSP, template1), global.bufptSP) # cropping to bbox
global.bufptR
```
```{r}
plot(global.bufptR, legend = T)
```
#### Save
```{r}
setwd(path.in.specific)
writeRaster(global.bufptR, paste0(spname, "_global.bboxR.img"), format = "HFA", overwrite = TRUE)
```

#### making FISHNET (assigning FNETID to each cell)

a dataframe version of the raster file for the buffered bounding box

```{r}
f1 <- sp::coordinates(global.bufboxR)
f2 <- cellFromXY(global.bufboxR, f1)
f3 <- as.data.frame(cbind(f1, f2))

# some dataframe clean-up
names(f3)[1:3] <- c("decimalLongitude", "decimalLatitude", "FNETID") # assign names
f3 <- f3[c("FNETID", "decimalLongitude", "decimalLatitude")] # reorder
class(f3)
head(f3)
```

##### double checks:

f2 value should eq. ncell in glboal.bufboxR below
```{r}
tail(f3, 1)
```
```{r}
ncell(global.bufboxR) # should eq. f2 value above
```
#### Set as spatial object

```{r}
global.fnetSF <- st_as_sf(f3, coords = c("decimalLongitude", "decimalLatitude"), crs = prj.wgs84, remove = F) # remove=F retains input x,y)
class(global.fnetSF)
head(global.fnetSF, 2) # examine
```
```{r}
setwd(path.in.specific)
#save
global.fnetDF <- st_drop_geometry(global.fnetSF) # build dataframe -  redundant, but whatever
write.csv(global.fnetDF, paste0(spname, "_global.fnetDF.csv"), row.names = FALSE)
head(global.fnetDF, 2)
```

### Create a Raster file of the FNET - this takes an hour
```{r eval=FALSE, include=FALSE}
global.fnetSP <- as_Spatial(global.fnetSF) # converts to spatial polygon class
global.fnetR <- crop(rasterize(global.fnetSP, template1), global.fnetSP)
global.fnetR
```

```{r eval=FALSE, include=FALSE}
setwd(path.in.general)
#save
writeRaster(global.fnetR, "global.fnetR.img", format = "HFA", overwrite = TRUE)
```

```{r}
setwd(path.in.general)
global.fnetR <- raster("global.fnetR.img")
global.fnetR
```

### Extract fishnet location points for each presence point
#### convert species presences x,y to spatial coordinate

```{r}
global.presDF <- st_drop_geometry(global.pres) # build dataframe -  redundant, but whatever
class(global.presDF)
```

```{r}
tru.xy <- global.presDF[c("decimalLongitude", "decimalLatitude")]  # get x,y of species presences
p1 <- sp::coordinates(tru.xy)
head(p1)
```

### Get FNETIDs that have spp presences

```{r}
#pres.fnetid <- cellFromXY(global.bufboxR, p1)
pres.fnetid <- cellFromXY(global.bufptR, p1) # FNETID of presence from pt-buffered poly
length(pres.fnetid)
head(pres.fnetid, 25) # FNETID numbers associated with species x,y
```

#### Check:
Make sure the species presences = the number of cells in original data
```{r}
length(pres.fnetid) == nrow(global.presDF)
```
### create species dataframe with FNETIDs
 (so that FNETIDs are linked to species presences)

```{r}
global.presFNET <- cbind(pres.fnetid, global.presDF) # add fishnet ID to your original dataframe
names(global.presFNET)[1] <- "FNETID" # name change 
head(global.presFNET, 5) # examine
```

### extract FNETID of modelling frame

```{r}
bufpt.fnetid <- raster::extract(global.bufptR, global.fnetDF[c("decimalLongitude", "decimalLatitude")])
bufpt.fnetid[1:6] # examine: NA
```
should be just NA and 1s
```{r}
unique(bufpt.fnetid)
```

### create modelling dataframe with FNETIDs

```{r}
global.bufptFNET <- cbind(global.fnetDF, bufpt.fnetid) # bind modelling frame w/FISHNET
head(global.bufptFNET, 2) # examine
```

##### Check...
```{r}
# some internal checking
length(global.bufptFNET$bufpt.fnetid) # number FNETIDs 
```
```{r}
ncell(global.bufptR) # should equal above
```
```{r}
table(global.bufptFNET$bufpt.fnetid)[[1]] # number of FNETIDs
```
```{r}
length(which(is.na(global.bufptFNET$bufpt.fnetid))) # NAS
```
latter # should be higher than former, indicating that there are plenty of cells within the buffer point modelling frame that don't have presence points....

```{r}
plot(global.bufboxR, legend = T)
plot(global.bufptR, col = "red", add = T, legend = F, main = "Points in modelling frame")  # main plot
points(global.presFNET$decimalLongitude, global.presFNET$decimalLatitude, pch = 20, col = "black") # add spp locations
```
### MERGE SPP LOCATION and MODELLING FRAME by FISHNET ID’S

```{r}
# examine dataframes
head(global.fnetDF, 2) # fishnet dataframe
```

```{r}
# begin merge: NOTE merge by=c("sort vars") & all.y=T options
m1 <- merge(global.fnetDF, global.bufptFNET, by = c("FNETID", "decimalLongitude", "decimalLatitude" ), all.y = T)
head(m1, 2) # examine: FNETID no longer ranked but not to worry ...
```
```{r}
nrow(m1)
```
```{r}
unique(m1$bufpt.fnetid)
```

```{r}
global.indexFNET <- merge(m1, global.presFNET, by = c("FNETID",  "decimalLongitude", "decimalLatitude"), all = T) # final merge:assign DF name
names(global.indexFNET)[4] <- "in.modFR" # change some names
names(global.indexFNET) # examine names
```
```{r}
head(global.indexFNET, 2) # final 
```
```{r}
unique(global.indexFNET$Species)
```


#### Save final fnet index dataframe

```{r}
setwd(path.in.specific)
save(global.indexFNET, file = paste0(spname, "_global.indexFNET.RData"))
  
# examine the fishnet
head(global.indexFNET)
```
```{r}
unique(global.indexFNET$in.modFR)
```


### EXTRACT PSEUDO-ABSENCES [X,Y]’s
### create conditional vectors: FNETIDs by spp locations & in modelling frame
```{r}
p1 <- global.indexFNET
p2.spp <- subset(p1$FNETID, p1$Species == 1) # FNETIDs of spp locations
p2.modFR <- subset(p1$FNETID, p1$in.modFR == 1) # FNETIDs in modelling frame
length(p2.spp) # should equal N of spp locations
```
```{r}
length(p2.modFR)
```
#### dropping presence cells

(leaving possible pseudo-absence cell FNETIDs)

```{r}
p3 <- p1[!p1$FNETID %in% p2.spp, ] # background from fishnet
p4 <- p1[!p1$FNETID %in% p2.spp & p1$FNETID %in% p2.modFR, ] # background from modelling frame
pseu.bufpt <- p4 # new name to dataframe: this dataframe used from now on
```


#### Drawing Pseudo-Absences

Note: The ulitmate # of pseudoabsences should ideally be somehwere between 2-3 X the # of presences.  However, if a portion of the buffered point area lands in the ocean,  then pseudo absences will retain "NA" values for environmental variables and will be excluded in the models.  Adjust the number below (currently at 2.5) to account for this issue.
```{r}
#   N=2*No. pres
set.seed(1234) # set seed to ensure repeatability
pseu.srs2 <- pseu.bufpt[sample(1:nrow(pseu.bufpt), 2.5 * table(p1$Species)[[1]], replace = F), ]
pseu.srs2$Species <- 0  # assign 0 to pseu.abs  
pseu.srs2$in.modFR <- 1  # assign 1 to in.modFR
dim(pseu.srs2) # dim[1] should equal N spp locations, dim[2] the No. variables
```
```{r}
head(pseu.srs2, 2) # examine
```

### MERGing PSEUDO-ABSENCES WITH TRUE PRESENCE DATAFRAME

```{r}
head(global.presFNET, 2)
```
```{r}
head(pseu.srs2, 2)# examine; both MUST have FNETID
```

#### merge with true presences

```{r}
global.PPsA <- merge(global.presFNET, pseu.srs2, by = c("FNETID", "Species", "decimalLongitude", "decimalLatitude"), all = T) # merge
global.PPsA$in.modFR <- NULL # drop in.modFR index no longer needed
dim(global.PPsA) # examine
```
```{r}
head(global.PPsA, 2) 
```

### create common x,y from tru pres and pseudo-abs x,y’s
```{r}
# create new vars wgs_x & wgs_y; used later in raster stack extraction
global.PPsA$tr.Lon <- ifelse(global.PPsA$Species == 0, global.PPsA$decimalLongitude, global.PPsA$decimalLongitude)
global.PPsA$tr.Lat <- ifelse(global.PPsA$Species == 0, global.PPsA$decimalLatitude, global.PPsA$decimalLatitude) 
global.PPsA[352:355, ] # examine a subset
```

### Saving

```{r}
global.PPsASF <- st_as_sf(global.PPsA, coords = c("tr.Lon", "tr.Lat"), crs = prj.wgs84, remove = FALSE)
class(global.PPsASF)
```
```{r}
global.PPsADF <- st_drop_geometry(global.PPsASF) # build dataframe
setwd(path.in.specific)
save("global.PPsADF", file = paste0(spname, "_global.PPsADF.RData"))
head(global.PPsADF, 2) # examine
```

## Point shapefile with geometry in R

```{r}
setwd(path.in.specific)
save("global.PPsASF", file = paste0(spname, "_global.PPsASF.RData"))
head(global.PPsASF, 2) # examine
```

Export as a point shapefile in ESRI format

```{r}
setwd(path.in.specific) # output path
st_write(global.PPsASF, dsn = ".", layer = paste0(spname, "_global_PPsASF"), driver = "ESRI Shapefile", delete_layer = T, delete_dsn = T) # output shapefile
```

## Get Global Environmental Data

Using WorldClim
(<a href="http://www.worldclim.org/" class="uri">http://www.worldclim.org/</a>)
bioclimatic dataset at resolution of 2.5 arcmin.

```{r}
# get WorldClim bioclimatic variable rasters
envs <- raster::getData(name = "worldclim", var = "bio", res = 2.5, lat = , lon = )
envs
```
```{r}
# get WorldClim bioclimatic variable rasters
elev <- raster::getData(name = "worldclim", var = "alt", res = 2.5, lat = , lon = )
elev
```
```{r}
combined <- raster::stack(envs, elev)
names(combined)
```

```{r}
setwd(path.in.general)
save(combined, file = "bioclimstack.RData") # save .RData
writeRaster(envs, "bioclimstack", format = "HFA", overwrite = T) # save as .img file
```
```{r}
setwd(path.in.general)
bioclim <- get(load("bioclimstack.RData"))
bioclim
```
```{r}
global.bufboxR
```
```{r}
bioclim[[1]]
```

### Extract predictor variables for a bioclim vars files using the presence:absence [X,Y]’s

##### First, get the presence-absence data

```{r}
pres.abs <- global.PPsADF
head(pres.abs, 2)
```
```{r}
table(pres.abs$Species) # examine frequencies
```
```{r}
t1 <- raster::extract(bioclim, pres.abs[, c("tr.Lon", "tr.Lat")]) # extract values from raster stack
head(t1, 30) # examine extracted matrix
```
There might be lots of NAs here if you buffered by a large distance so that the pseudo abs fall in the ocean. (adjust above if problematic)


### Create a new dataframe by binding the extracted predictor variable values to the pres abs dataframe

```{r}
global.trTOPO <- cbind(pres.abs, t1) # bind to train dataframe
  head(global.trTOPO, 5) # examine training data frame
```
```{r}
nrow(pres.abs)
```
```{r}
dim(t1)
```

```{r}
# examine extractions
nrow(subset(global.trTOPO, is.na(as.character(global.trTOPO$bio1))))
```
How many NA's exist? preferably very few

```{r}
global.trTOPO
```

```{r}
global.trTOPOSF <- st_as_sf(global.trTOPO, coords = c("tr.Lon", "tr.Lat"), crs = prj.wgs84, remove = FALSE)
class(global.trTOPOSF)
```


```{r}
zeros <- subset(global.trTOPOSF, Species == 0)
ones <- subset(global.trTOPOSF, Species == 1)


plot(st_geometry(zeros), axes = T,col = "blue")
plot(st_geometry(ones), add = T,  col = "red")
```
```{r}
nrow(zeros)
```
```{r}
nrow(ones)
```


##### Saving
```{r}
setwd(path.in.specific)
write.csv(global.trTOPO, file = paste0(spname, "_global_trTOPO.csv"), row.names = F) # save .csv
save(global.trTOPO, file = "_global.trTOPO.RData") # save .RData
st_write(global.trTOPOSF, dsn = ".", layer = paste0(spname, "_global.trTOPOSF"), driver = "ESRI Shapefile", delete_layer = T, delete_dsn = T) # output shapefile
```

## EXAMINE ENVIRONMENTAL RELATIONSHIPS


```{r}
setwd(path.in.specific)
tr.data <- read.csv( file = paste0(spname, "_global_trTOPO.csv"))
tr.data <- subset(tr.data, !is.na(tr.data$bio1))
dim(tr.data)
```
```{r}
table(tr.data$Species)
```

#### Description of World BIOCLIM environmental variables
    BIO1 = Annual Mean Temperature
    BIO2 = Mean Diurnal Range (Mean of monthly (max temp - min temp))
    BIO3 = Isothermality (BIO2/BIO7) (×100)
    BIO4 = Temperature Seasonality (standard deviation ×100)
    BIO5 = Max Temperature of Warmest Month
    BIO6 = Min Temperature of Coldest Month
    BIO7 = Temperature Annual Range (BIO5-BIO6)
    BIO8 = Mean Temperature of Wettest Quarter
    BIO9 = Mean Temperature of Driest Quarter
    BIO10 = Mean Temperature of Warmest Quarter
    BIO11 = Mean Temperature of Coldest Quarter
    BIO12 = Annual Precipitation
    BIO13 = Precipitation of Wettest Month
    BIO14 = Precipitation of Driest Month
    BIO15 = Precipitation Seasonality (Coefficient of Variation)
    BIO16 = Precipitation of Wettest Quarter
    BIO17 = Precipitation of Driest Quarter
    BIO18 = Precipitation of Warmest Quarter
    BIO19 = Precipitation of Coldest Quarter

```{r}
tr.data
```


```{r}
occurrences <- c("FNETID", "Species", "tr.Lon", "tr.Lat")

tempvars <- c('bio1', 'bio2', 'bio3', 'bio4', 'bio5', 'bio6', 'bio7', 'bio8', 'bio9', 'bio10', 'bio11' )
precipvars <- c('bio12', 'bio13', 'bio14', 'bio15', 'bio16', 'bio17', 'bio18', 'bio19', 'alt')

allpreds <- c(tempvars, precipvars)
```

```{r}
tr.temp <- cbind(tr.data %>% dplyr::select(all_of(occurrences)), tr.data %>% dplyr::select(all_of(tempvars)))
tr.temp$pred_type <- "temp"
```
```{r}
tr.prec <- cbind(tr.data %>% dplyr::select(all_of(occurrences)), tr.data %>% dplyr::select(all_of(precipvars)))
tr.prec$pred_type <- "precip"
```
Creating an empty dataframe to merge together

```{r}
tr.DFgrouped <-  setNames(data.frame(matrix(ncol = length(allpreds), nrow = 0)), allpreds)
```

```{r}
tr.DFgrouped <- dplyr::bind_rows(mutate_all(tr.DFgrouped, as.character), mutate_all(tr.temp, as.character), mutate_all(tr.prec, as.character))
head(tr.DFgrouped)
```

### CORRELATIONS AMONG ALL PREDICTORS
```{r}
tr.data
```


```{r}
# numeric correlations
#Using spearman because it has no assumption of normality
  cut.point <- 0.7 # set cutpoint for correlation
  c1 <- cor(tr.data[, c(9:length(tr.data))], use = "pairwise.complete.obs", method = "spearman") # est. correlation
```
```{r}
c2 <- subset(c1 > cut.point | c1 < -cut.point) # matrix of cor>cutpoint
c2 # examine; FALSE indicates cor<cutpoint 
```
How many correlations are there?
```{r}
table(c2)[[2]] - 19
```
START MODIFIED panel.cor CORRELATION FUNCTION
```{r}
#   determine correlations among predictor variables: modified from 
#   http://addictedtor.free.fr/graphiques/graphcode.php?graph=137
panel.cor <- function(x, y, digits=2, prefix="", cex.cor) 
  { usr <- par("usr"); on.exit(par(usr)) 
    par(usr = c(0, 1, 0, 1)) 
    r <- abs(cor(x, y, use = "pairwise.complete.obs")) 
    txt <- format(c(r, 0.123456789), digits=digits)[1] 
    txt <- paste(prefix, txt, sep="") 
    if(missing(cex.cor)) cex <- 0.8/strwidth(txt) 

    test <- cor.test(x,y) 
    # borrowed from printCoefmat
    Signif <- symnum(test$p.value, corr = FALSE, na = FALSE, 
                  cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("***", "**", "*", ".", " ")) 

    text(0.5, 0.5, txt, cex = cex * r)
    text(.8, .8, Signif, cex=cex, col=2) 
}
# END MODIFIED panel.cor CORRELATION FUNCTION
```

### CORRELATIONS BETWEEN TEMPERATURE PREDICTORS
```{r}
# numeric correlations
#Using spearman because it has no assumption of normality
cut.point <- 0.7 # set cutpoint for correlation
c1 <- cor(tr.temp[, c(5:(ncol(tr.temp)-1))], use = "pairwise.complete.obs",method = "spearman") # est. correlation
c2 <- subset(c1 > cut.point | c1 < -cut.point) # matrix of cor>cutpoint
```
```{r}
# plot correlations using modified panel.cor function
  pairs(tr.temp[, c(5:(ncol(tr.temp)-1))], lower.panel = panel.smooth, 
    upper.panel = panel.cor, main = "Temp Variables")
```
START VARIABLE IMPORTANCE AMONG TEMPERATURE PREDICTORS
Modified from: Niklaus E. Zimmermann, WSL / ETH Zurich
```{r}
#### START function variable importance
varimp.glm <- function(tr.spp, tr.var, pres, pf, pl) {
  tmp.mat <- matrix(ncol = 2, nrow = (pl - pf + 1))
  for (i in pf:pl) {
    # option: linear+quadratic; linear only
    tmp <- glm(tr.spp[, pres] ~ tr.var[, i] + I((tr.var[, i])^2), na.action = na.omit, 
      family = binomial)
    # linear only glm
    #tmp <- glm(tr.spp[, pres] ~ tr.var[, i], na.action = na.omit, family = binomial)
    tmp.mat[(i - pf + 1), 1] <- tmp$aic
    tmp.mat[(i - pf + 1), 2] <- (1 - (tmp$deviance/tmp$null.deviance))
    }
  return(tmp.mat)
  } 
#### END function variable importance
```

```{r}
# estimate VIP values => AIC & Adj deviance
  tr.vip <- tr.temp[, c(2, 5:(ncol(tr.temp)-1))] # keep only P/A & predictors
  pres <- 1 # column for presence:absence
  v.start <- 2 # column start predictor variables
  v.stop <- ncol(tr.vip) # last column predictor variables
  v.num <- v.stop - 1 # number predictor variables
  dev.fit <- varimp.glm(tr.vip, tr.vip, pres, v.start, v.stop) # call VIP function
  dev.fit # output matrix; col=1 AIC, col=2 Adj deviance
```
```{r}
# built basic barplot if desired
  d.max <- ceiling(signif(max(dev.fit[, 2]), 2) * 10)/10 # max of y-axis
  ylim.r <- range(0, d.max) # range y-axis
  x.labs <- names(tr.vip[2:v.stop]) # x-axis labels
  barplot(dev.fit[, 2], col = "darkgreen", ylim = ylim.r, main = "temp VIPs", 
    ylab = "adj.D2", names = x.labs) # barplot
  abline(h = 0) # add horizontal line
  abline(mean(dev.fit[, 2]), 0, lt = 3) # ref lines; dash=mean adj.dev 
```

### CORRELATIONS BETWEEN PRECIPITATION PREDICTORS
```{r}
# numeric correlations
#Using spearman because it has no assumption of normality
cut.point <- 0.7 # set cutpoint for correlation
c1 <- cor(tr.prec[, c(5:(ncol(tr.prec)-1))], use = "pairwise.complete.obs",method = "spearman") # est. correlation
c2 <- subset(c1 > cut.point | c1 < -cut.point) # matrix of cor>cutpoint
c2 # examine; FALSE indicates cor<cutpoint 
```
```{r}
# plot correlations using modified panel.cor function
  pairs(tr.prec[, c(5:(ncol(tr.prec)-1))], lower.panel = panel.smooth, 
    upper.panel = panel.cor, main = "Precip Variables")
```
```{r}
# estimate VIP values => AIC & Adj deviance
  tr.vip <- tr.prec[, c(2, 5:(ncol(tr.prec)-1))] # keep only P/A & predictors
  pres <- 1 # column for presence:absence
  v.start <- 2 # column start predictor variables
  v.stop <- ncol(tr.vip) # last column predictor variables
  v.num <- v.stop - 1 # number predictor variables
  dev.fit <- varimp.glm(tr.vip, tr.vip, pres, v.start, v.stop) # call VIP function
  dev.fit # output matrix; col=1 AIC, col=2 Adj deviance
```
```{r}
# built basic barplot if desired
  d.max <- ceiling(signif(max(dev.fit[, 2]), 2) * 10)/10 # max of y-axis
  ylim.r <- range(0, d.max) # range y-axis
  x.labs <- names(tr.vip[2:v.stop]) # x-axis labels
  barplot(dev.fit[, 2], col = "darkgreen", ylim = ylim.r, main = "precip VIPs", 
    ylab = "adj.D2", names = x.labs) # barplot
  abline(h = 0) # add horizontal line
  abline(mean(dev.fit[, 2]), 0, lt = 3) # ref lines; dash=mean adj.dev 
```

#### Which variables would you like to keep?
```{r}
keep <- c("bio2", "bio5", "bio8","bio11", "bio12", "bio15",  "bio18", "alt")
tr.picked <- cbind(tr.data %>% dplyr::select(all_of(occurrences)), tr.data %>% dplyr::select(all_of(keep)))
head(tr.picked)
```
```{r}
par(mfrow = c(1, 5))
boxplot(tr.picked[,length(tr.picked) - length(keep) ]~ tr.picked$Species, xlab = "Presence:Absence", 
    ylab = keep[1])

boxplot(tr.picked[,length(tr.picked) - length(keep) + 1 ] ~ tr.picked$Species, xlab = "Presence:Absence",
    ylab = keep[2])

boxplot(tr.picked[,length(tr.picked) - length(keep) + 2 ] ~ tr.picked$Species, xlab = "Presence:Absence",
    ylab = keep[3])

boxplot(tr.picked[,length(tr.picked) - length(keep) + 3 ] ~ tr.picked$Species, xlab = "Presence:Absence",
    ylab = keep[4])

boxplot(tr.picked[,length(tr.picked) - length(keep) + 4 ] ~ tr.picked$Species, xlab = "Presence:Absence",
    ylab = keep[5])

```

```{r}
tr.picked %>% 
  pivot_longer(names_to = "covariate", values_to = "value", - (FNETID:tr.Lat)) %>% 
  ggplot() +
  geom_freqpoly(aes(x = value, y = ..density.., color = factor(Species))) +
  facet_wrap(~covariate, scales = "free") +
  theme_minimal()
```


### Saving

```{r}
setwd(path.in.specific)
save(keep, file = paste0( spname, "_list_picked_vars_global.RData")) # save .RData
```

```{r}
setwd(path.in.specific)
write.csv(tr.picked, file = paste0( spname, "_tr_global_pickedvars.csv"), row.names = F) # save .csv
save(tr.picked, file = paste0( spname, "_tr_global_pickedvars.RData")) # save .RData

tr_globalSF <- st_as_sf(tr.picked, coords = c("tr.Lon", "tr.Lat"), crs = prj.wgs84, remove = FALSE)
save(tr_globalSF, file = paste0( spname, "_tr_globalSF_pickedvars.RData")) # save .RData
st_write(tr_globalSF, dsn = ".", layer = paste0( spname, "_tr_globalSF_pickedvars"), driver = "ESRI Shapefile", delete_layer = T, delete_dsn = T) #
```

```{r}
setwd(path.in.specific)
write.csv(tr.data, file = paste0( spname, "_tr_global_allvars.csv"), row.names = F) # save .csv
save(tr.data, file = paste0( spname, "_tr_global_allvars.RData")) # save .RData

tr_globalSF <- st_as_sf(tr.data, coords = c("tr.Lon", "tr.Lat"), crs = prj.wgs84, remove = FALSE)
save(tr_globalSF, file = paste0( spname, "_tr_globalSF_allvars.RData")) # save .RData
st_write(tr_globalSF, dsn = ".", layer = paste0( spname, "_tr_globalSF_allvars"), driver = "ESRI Shapefile", delete_layer = T, delete_dsn = T) # 
```

```{r}
beep(sound = 2)
beep(sound = 2)
beep(sound = 8)
beep(sound = 4)
```
